---
title: "Dates"
author: "Bakro"
date: "10/5/2021"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r  setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#klippy::klippy()
```

```{r echo = FALSE , message= FALSE}
library(tidyverse)
library(lubridate)
library(nycflights13)
dashes <- function()
{
  cat('\n',"------------------------------------",'\n')
}
```

A **date-time** is a point on the timeline,stored as the number of
seconds since

1970-01-01 00:00:00 UTC

"2017-11-28 12:00:00 UTC"

**today & now**

```{r}
dt <- as_datetime(1511870400)
dt

cat(paste("today : " ,today()));
print(paste("now : " , now()));  
```

# PARSE DATE-TIMES

> Convert strings or numbers to date-times

1\. Identify the order of the year (y), month (m), day (d), hour (h),

minute (m) and second (s) elements in your data.

2\. Use the function below whose name replicates the order. Each

accepts a tz argument to set the time zone, e.g. ymd(x, tz = "UTC").

```{r}

cat(paste("ymd : 2017-01-31 -->" , ymd("2017-01-31"))) ;
cat(paste("mdy : January 31st, 2017 -->" , mdy("January 31st, 2017")));
cat(paste("dmy(31-Jan-2017) -->" , dmy("31-Jan-2017"))); 
cat(paste("ymd(20170131) -->" , ymd(20170131)));

```

```{r}
datechr <- "2017-11-28 12:00:00 UTC"
cat(paste("ymd hms :" ,ymd_hms("2017-01-31 20:11:59") )) ; 
cat(paste("mdy hm :" , mdy_hm("01/31/2017 08:01") )) ;
cat(paste("ymd" ,ymd(20170131, tz = "UTC")));
```



mdy_hms(), mdy_hm(), mdy_h().

mdy_hms("11/28/2017 1:02:03")


dmy_hms(), dmy_hm(), dmy_h().

dmy_hms("1 Jan 2017 23:59:59")


ymd(), ydm(). ymd(20170131)


mdy(), myd(). mdy("July 4th, 2000")


dmy(), dym(). dmy("4th of July '99")


yq() Q for quarter. yq("2001: Q3")


my(), ym(). my("07-2020")




# make_datetime

```{r}
flights_dt  <- flights %>% 
  select(year, month, day, hour, minute) %>% 
  mutate(departure = make_datetime(year, month, day, hour, minute))
head(flights_dt)
```

# Date-time components

```{r}
datetime <- ymd_hms("2016-07-08 12:34:56")
date(datetime)   # "2016-07-08"
isoyear(datetime)  #2016
year(datetime)    #2016
month(datetime)   #7
mday(datetime)    #8
yday(datetime)    #190 order day in  year
wday(datetime, label  =TRUE) # fr order day in week
weekdays(datetime )  # Freitag"
hour(datetime)    #12
minute(datetime)  # 34
second(datetime)    # 56
tz(datetime)        # "UTC"
quarter(datetime)   #3
semester(datetime)  #2
am(datetime)    # false
pm(datetime)  # True


```

# update dates

```{r}
datetime
update(datetime,mday = 2 , hour = 1)
```

# Round Date-times(optional)

```{r}
floor_date(datetime, unit = "second")
#Round down to nearest unit.
floor_date(datetime, unit = "month")
round_date(datetime, unit = "second")
#Round to nearest unit.
round_date(datetime, unit = "month")
ceiling_date(datetime, unit = "second",
change_on_boundary = NULL)
#Round up to nearest unit.
ceiling_date(datetime, unit = "month")
```

# Time Zones (optional)

> R recognizes \~600 time zones.
>
> Each encodes the time zone, Daylight Savings Time, and historical
> calendar variations for an area.
>
> R assignsone time zone per vector.
>
> Use the UTC time zone to avoid Daylight Savings.
>
> OlsonNames() Returns a list of valid time zone names. OlsonNames()
>
> Sys.timezone() Gets current time zone.

-   the same date-time in a new time zone (a new clock time).

Also local_time(dt, tz, units).

-   force_tz(time, tzone = "")

    Get the same clock time in a new time zone (a new date-time).

    Also force_tzs().

```{r}

with_tz(datetime, "US/Pacific")

force_tz(datetime, "US/Pacific")
```

# Math with Date_times 

## periods

Add or subtract periods to model events that happen at specific clock

times, like the NYSE opening bell.

```{r}
p <- years(1) + months(3) + weeks(2) + hours(12)+ minutes(3) + 
    seconds(3) + milliseconds(20)
p
as.period(p)
period_to_seconds(p)
```

## duration

Add or subtract durations to model physical processes, like battery
life.

Durations are stored as seconds, the only time unit with a consistent
length.

Difftimes are a class of durations found in base R.

```{r}
dyears(x = 1) #31536000x seconds.
dmonths(x = 1) #2629800x seconds.
dweeks(x = 1) #604800x seconds.
ddays(x = 1) #86400x seconds.
dhours(x = 1)# 3600x seconds.
dminutes(x = 1)# 60x seconds.
dseconds(x = 1) #x seconds.
dmilliseconds(x = 1)# x x 10-3 seconds.
dmicroseconds(x = 1) #x x 10-6 seconds.
dnanoseconds(x = 1) #x x 10-9 seconds.
dpicoseconds(x = 1) #x x 10-12 seconds.

```

## intervals

Divide an interval by a duration to determine its physical length,
divide

an interval by a period to determine its implied length in clock time.

```{r}
i <- interval( datetime,ymd("2017-01-01")) 
j <- datetime %--% ymd("2017-12-31")
datetime
i
j
```

a %within% b Does interval or date-time a fall

within interval b?

```{r}
 now() %within% i
```

int_start(int) Access/set the start date-time of

an interval. Also int_end().

```{r}
int_start(i) <- now();
int_start(i)
```

int_aligns(int1, int2) Do two intervals share a

boundary? Also int_overlaps().

```{r}
 int_aligns(i, j)
```

int_diff(times) Make the intervals that occur

between the date-times in a vector.

```{r}
datetime -> dt
v <-c(dt, dt + 100, dt + 1000); int_diff(v)
```

int_flip(int) Reverse the direction of an

interval. Also int_standardize().

```{r}
 int_flip(i)
```

int_length(int) Length in seconds.

```{r}
int_length(i)
```

int_shift(int, by) Shifts an interval up or down

the timeline by a timespan.

```{r}
int_shift(i, days(-1))
```

as.interval(x, start, ...) Coerce a timespan to

an interval with the start date-time. Also

is.interval().

```{r}
as.interval(days(1), start = now())
```


sequence of time 
```{r}
start <- ymd("2016-07-08 ")
end <- ymd("2016-07-12 ")
seq(start , end , by =1)
```

