---
title: "Strings"
author: "Bakro"
date: "10/5/2021"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r  setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basics

```{r  }
library(stringr)

```

```{r}
firstname <- "abd" 
lastname <- "alex"
 cat(firstname, "\n", lastname, sep = "")
```

```{r}
vec <- c(firstname,lastname)
length(vec)
summary(vec)
attributes(vec)
```

```{r}
vcNum <- formatC(c(100.23,2400.59), digits = 3, width = 4, flag = "0"); vcNum
```

The **stringr** package provides a set of internally consistent tools
for working with character strings, i.e. sequences of characters
surrounded by quotation marks.

# Detect Matches

## str_detect()

str_detect(string, pattern, negate = FALSE)

Detect the presence of a pattern match in astring. Also str_like().

```{r}
fruit <- c("apple","orange","kiwi","pear" ,"grapes")
ind <- str_detect(fruit, "a")
fruit[ind]
```

## str_starts()

str_starts(string, pattern, negate = FALSE)

Detect the presence of a pattern match at

the beginning of a string. Also str_ends().

```{r}
str_starts(fruit, "a")
```

## str_which()

str_which(string, pattern, negate = FALSE)

Find the indexes of strings that contain

a pattern match.

```{r}
str_which(fruit, "a")
```

## str_locate()

str_locate(string, pattern) Locate the

positions of pattern matches in a string.

Also **str_locate_all().**

```{r}
str_locate(fruit, "a")
```

## str_count()

str_count(string, pattern) Count the number

of matches in a string. str_count(fruit, "a")

```{r}
fruit[6] <- "aa"
 str_count(fruit, "a")
```

# Mutate Strings

## str_sub()

str_sub() \<- value. Replace substrings byidentifying the substrings
with str_sub() and

assigning into the results.

```{r}
fruitcopy <- c("apple","orange","kiwi","pear" ,"grapes")
str_sub(fruitcopy, 1, 3)[] <- "str"  # add indecies 
fruitcopy
```

## str_replace()

str_replace(string, pattern, replacement)

Replace the first matched pattern in each

string. Also **str_remove()**.

```{r}
str_replace(fruit,"p","-") 
str_replace(fruit,"p","")  # same str_remove
str_remove(fruit,"p") 
```

## str_replace_all()

str_replace_all(string, pattern, replacement)

Replace all matched patterns in each string.

Also **str_remove_all()**.

```{r}
str_replace_all(fruit, "p", "-")
```

## str_to_lower()

str_to_lower(string, locale = "en")1

Convert strings to lower case.

```{r}
str_to_lower(c("Fxy","YxZ"))
```

## str_to_upper()

str_to_upper(string, locale = "en")1

Convert strings to upper case.

```{r}
str_to_upper(c("Fxy","YxZ"))
```

## str_to_title()

str_to_title(string, locale = "en")1 Convert

strings to title case. Also str_to_sentence().

```{r}
str_to_title(c("Fxy","YxZ"))
```

## reverse string

```{r}
start <- proc.time()
t <- 0L
for (i in c(1L:10000L)) {
  t <- t + i
}
end <- proc.time()
print(t)
print(end - start)
```

# Subset Strings

## str_sub()

str_sub(string, start = 1L, end = -1L) Extract

substrings from a character vector.

```{r}
fruit
str_sub(fruit, 1, 3)
str_sub(fruit, -2)
```

## str_subset()

str_subset(string, pattern, negate = FALSE)

Return only the strings that contain a pattern match.

```{r}
str_subset(fruit, "p")
```

## str_extract()

str_extract(string, pattern) Return the first

pattern match found in each string, as a vector.

Also str_extract_all() to return every pattern match.

```{r}
 str_extract(fruit, "[aeiou]")
```

## str_match()

str_match(string, pattern) Return the

first pattern match found in each string, as

a matrix with a column for each ( ) group in

pattern. Also str_match_all().

```{r}
str_match(sentences[1:5], "(a|the) ([^ +])")
```

# Join and Split

## str_c()

str_c(..., sep = "", collapse = NULL) Join

multiple strings into a single string.

```{r}
str_c(letters, LETTERS)
```

## str_flatten()

str_flatten(string, collapse = "") Combines

into a single string, separated by collapse.

```{r}
str_flatten(fruit, ", ")
```

## str_dup()

str_dup(string, times) Repeat strings times

times. Also str_unique() to remove duplicates.

```{r}
str_dup(fruit, times = 2)
```

## str_split_fixed()

str_split_fixed(string, pattern, n) Split a

vector of strings into a matrix of substrings

(splitting at occurrences of a pattern match).

Also **str_split()** to return a list of substrings

and **str_split_n()** to return the nth substring.

```{r}
str_split_fixed(sentences[1:3], " ", n=3) #tokens max =3
str_split_fixed(sentences[1:3], " ", n= Inf) # all tokens 
```

## str_glue()

str_glue(..., .sep = "", .envir = parent.frame())

Create a string from strings and {expressions}

to evaluate.

```{r}
str_glue("Pi is {pi}")
```

## str_glue_data()

str_glue_data(.x, ..., .sep = "", .envir =

parent.frame(), .na = "NA") Use a data frame,

list, or environment to create a string from

strings and {expressions} to evaluate.

```{r}
t<-str_glue_data(mtcars, "{rownames(mtcars)} has

{hp} hp")
t[1:5]
```

# Manage Lengths

## str_length()

str_length(string) The width of strings (i.e.

number of code points, which generally equals

the number of characters).

 

```{r}
 str_length(fruit)
```

## str_pad()

str_pad(string, width, side = c("left", "right",

"both"), pad = " ") Pad strings to constant

width.

```{r}
str_pad(fruit, 17)
```

## str_trunc()

str_trunc(string, width, side = c("right", "left",

"center"), ellipsis = "...") Truncate the width

of strings, replacing content with ellipsis.

```{r}
str_trunc(sentences[1:20], 6)
```

## str_trim()

str_trim(string, side = c("both", "left", "right"))

Trim whitespace from the start and/or end of

a string.

```{r}
str_trim(str_pad(fruit, 17))
```

## str_squish()

str_squish(string) Trim whitespace from each

end and collapse multiple spaces into single spaces.

```{r}
str_squish(str_pad(fruit, 17, "both"))
```

# Order Strings

## str_order()

str_order(x, decreasing = FALSE, na_last =

TRUE, locale = "en", numeric = FALSE, ...)1

Return the vector of indexes that sorts a

character vector.

```{r}
fruit[str_order(fruit)]
```

## str_sort()

str_sort(x, decreasing = FALSE, na_last =

TRUE, locale = "en", numeric = FALSE, ...)1

Sort a character vector.

```{r}
str_sort(fruit,decreasing = TRUE)
```

# Helpers

## str_conv()

str_conv(string, encoding) Override the

encoding of a string.

```{r}
str_conv(fruit,"ISO-8859-1")
```

## str_view_all()

str_view_all(string, pattern, match = NA)

View HTML rendering of all regex matches.

Also **str_view()** to see only the first match.

```{r}
str_view_all(sentences[1:5], "[aeiou]")
```


