---
title: "programming"
author: "Abdulmajid Bakro"
date: "10/5/2021"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sqldf)
```

# Conditions

## variable

```{r}
num1 <- 13 
num2 <- 20
if (num1 > num2)
{rint("number1 is bigger than number2")
  } else 
{ print("number1 is smaller than number2")}


x <- 0
if (x < 0) {
print("Negative number")
} else if (x > 0) {
print("Positive number")
} else
print("Zero")
```

## vector

```{r}
vec1  <- c(1.1, 2.2, 3.3)
vec2 <- c(1.1, 2.2, 3.5)
if (isTRUE(all.equal(vec1, vec2))) {
   cat("same vector \n")
}

```

-   isTRUE(), isFALSE()

-   is.logical(), is.integer(), is.numeric(), is.double(),
    is.character(),

-   is.vector(), is.factor(), is.matirx(), is.array(), is.data.frame(),
    is.list()

-   is.na(), is.name(), is.finite(), is.infinite()

-   is.expression,

-   is.function().

## switch()-Statement

```{r}

switch(1,"red","green","blue")

vnVal <- 2                          
switch(vnVal,    # muss return position                   
       print("Val is 1"),           
       print("Val is 2"),          
       print("Val is 3"))
```

# LOOPS

you can do a work for n times for example you want to draw all graphs
(hist for contentious ,pie for categorical ) that save time and you can
use that for any data set so that you can save it as function!!

## For

```{r}
sum <- 0                                 
for(ind in 1:5) {
    if(ind == 3) next  # skip this value from sum 
  sum <- sum + ind                 
                           
  cat("iterrate ", ind, "\n",                
      "cum_sum ", sum, "\n")  
} 
```

## loops and vectors

difference between them in performance !!

```{r}
vec1 <- c(1:1000000)   #  initialize is easy !!
vec2 <- rep(c(1:1000),1000)

start <- proc.time()
 t<- vec1*vec2
end <- proc.time()
print(end - start)



start <- proc.time()
v <- NULL ;
for (ind in 1:1000000) {
  v[ind] = vec1[ind]*vec2[ind]
}
end <- proc.time()
print(end - start)
```

## while

```{r}
ind <- 1          
arr <- NULL   

while(ind <= 10) {                     
   arr[ind] <- 2 * ind + 1 
    if(arr[ind] > 20) break
   ind <- ind + 1  # iterate or var condition 
} 
arr
```

## repeat

```{r}
x <- 1
repeat{
  
   if (x> 10) break ;
   x <- x + 1 
  
}
```

## apply

-   less typing

-   readable code

-   not faster than for loop

apply(X, MARGIN, FUN, ...)

-   X: **Input-Array**

-   MARGIN: 1 - rows, 2 - cols, c(1, 2) - rows & cols

-   FUN: any standard function or your function

```{r}
mat <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), nrow = 3, ncol =3)
mat
rMeans <- NULL; cMeans <- NULL
for(vn in 1:nrow(mat)) { 
   rMeans[vn] <- mean(mat[vn,]) 
}
rMeans

for(vn in 1:ncol(mat)) { 
   cMeans[vn] <- mean(mat[,vn])
} 
cMeans

```

```{r}
rMeans <- apply(mat, 1, mean)  # data,when want row (1) ,function  
rMeans

cMeans <- apply(mat, 2, mean) # data,when want col (2) ,function 
cMeans
cMeans <- apply(mat, c(1,2), mean) # rows,cols
cMeans
```

```{r}
x <- matrix(rnorm(200),20,10)
apply(x,1,quantile,props = c(0.25,0.75))
```

## tapply

vecors , matrix,...

```{r}


tapply(diamonds$price, diamonds$cut, mean)
tapply(diamonds$price, diamonds$cut, range)
```

## lapply & sapply

**apply on list**

```{r}
mylist <- list(vnEins = seq(1, 10, by = 0.3), 
              vnZwei = seq(100, 150, by = 2), 
              vnDrei = seq(-10, -5, by = 1), 
              vnVier = seq(30, 45, by = 1.3))
mylist

lapply(mylist, mean)
sapply(mylist, mean)
```

## mapply

**apply on list / apply for multiple lists parallel over set of
argument**

```{r}
mapply(rep,1:4,4:1)
#noise(1,1,2)
#mapply(plot,1:5,1:5,2)
# noise(1,1,2) ,noise(2,2,2)...noise(5,5,2)

```

## split

x: vector,dataframe

f: factor

drop : drop factor null group

```{r}
x <- c(rnorm(10),runif(10),rnorm(10))
f <- gl(3,10)
split(x,f)
```

# Function

when you write summary (data) you show info about data

Q : what you will do if this function not found ?

-   you have used a lot of function that save a lot of time

-   (divide and conquer ) big to small

-   reusable functions over and over

as we will see in explore function

```{r}
# function that sum only number and remove strings with show warning 
## if you have multi variable you can use vector or list 
sumnumbers <- function(a,b)
{
  result <- a + b 
  result
  
}

sumnumbers(3,4)
#summnumbers(3,4,5)  error
#sumnumbers(3,"b")   error

```

**on.exit ()**: Command sequence that is always used when the function
is exited is performed **warning ()**: Output of a self-created warning
**stop ():** Abort of the function with self-created stop message
**stopifnot ():** Terminate the function if a condition is met! **try
():** handling of errors for a statement **tryCatch ():** Handling of
errors, warnings, messages, aborts

## Function Arguments And Their Default

```{r}
MyThirdFun <- function(n, y = 2) 
{
  # Compute the power of n to the y
  n^y  
}

# Specify both args
MyThirdFun(2,3) 

# Just specify the first arg
MyThirdFun(2)
```

```{r}
MyThirdFun <- function(n, y = seq(0.05, 0.1, by = 0.01))
{
  # Compute the power of `n` to the `y`
  n^y  
}

# As before, specify both args
MyThirdFun(2,3)

# Compute all possible according to given default
MyThirdFun(2)  

# Specify no arguments: error!
# MyThirdFun()   
```

## missing

```{r}
# Equivalent alternative:
MyFourthFun <- function(n, y)  
 { if(missing(y))
   {
    y <- seq(0.05, 0.1, by = 0.01)
 }
 return(n^y)
}

MyFourthFun(2,3)

# Compute all possible according to given default
MyFourthFun(2)  

# Specify no argument: error!
# MyFourthFun() 
```

```{r}
MyFourthFun <- function(n, y)  
{  
  # Uncomment `print()` calls to check passed values
  # print(n)
  # print(y)
  if(missing(n)) n=2;
  if(missing(y)) y=0.05;
  if(!y %in% seq(0.05, 1, by = 0.02)) print("value must be <= 1)")
  else return(n^y)
}

# Calculation will be carried out
MyFourthFun(2,0.07)

# Prints an error, `y` is not in the allowed list 
MyFourthFun(2,3)

# Use `y` default
MyFourthFun(2)   

# No arguments: both `n` and `y` defaults are used
MyFourthFun()     
```

## the "..." Argument

```{r}
args(paste)
args(cat)

paste("a" ,"b",sep = ":")
paste("a" ,"b",se = ":")
```

## Scoping Rules - Symbol Binding

```{r}
search()  # search for fun depend on this order
# when you use function take local but not in sub functions  
y <- 10
f<- function(x){
      y <- 2
      y^2 + g(x)
}
g <- function(x){
      x*y
}

f(3)
# x = 3 y = 2    2
```

trycatch or if to predict code (when use try catch )

```{r}
div <-
  function(a, b) {
    objRes <- tryCatch(
      { 
        a/b
      },
      error = function(e) {paste("error: ", e)},
      warning = function(w) {paste("Warnung : ", w)}, 
      finally = { cat("finally \n") }
    )
    cat("result: ", objRes, "\n")
    return(objRes)
  }
t <- div(1,0)
t
```

# PIPES
it sequence of instruction that implement by their order without any 
saving
***remember to add this to dplyr***

for different groups what is mean of attentation and Dv for people with 
iq between 90 and 110

```{r}
load(url("http://dwoll.de/rKompakt/dat.Rdata"))

str(datW)

datW_MW2 <- 
   datW %>% 
   dplyr::group_by(group) %>%
   dplyr::select(group, iq, attention, DV) %>%
   dplyr::filter((iq >= 90) & (iq <= 110)) %>%
   dplyr::summarise(att_m = mean(attention, na.rm = TRUE), 
                    dv_m = mean(DV, na.rm = TRUE))
```

using selection

```{r}
# think group a key word not variable 
 sqldf("select avg(DV),avg(attention)  from datW where iq between 90 and 110 group by hair")
```

## pipes vs sqldf

```{r}
start <- proc.time()
#datW3 <- rename(datW ,group = group1)
# think group a key word not variable 
 sqldf("select avg(DV),avg(attention)  from datW where iq between 90 and 110 group by hair")
end <- proc.time()
print(end - start)


print("--------------------------------")

start <- proc.time()

 datW %>% 
   dplyr::group_by(hair) %>%
   dplyr::select(hair, iq, attention, DV) %>%
   dplyr::filter((iq >= 90) & (iq <= 110)) %>%
   dplyr::summarise(att_m = mean(attention, na.rm = TRUE), 
                    dv_m = mean(DV, na.rm = TRUE))
end <- proc.time()
print(end - start)
```

# Debugging

message/error/warning/logical error

```{r}
log(-1)  # warning
```

```{r}
printmessage <- function(x){
  if(x<0)
    print("neg")
  else
    print("pos")
  invisible(x)
}
printmessage(1)
#printmessage(NA)
```

```{r}
printmessage2 <- function(x){
  if(is.na(x))
    print("not available")
  else if(x<0)
    print("neg")
  else
    print("pos")
  invisible(x)
}
printmessage2(1)
printmessage2(NA)
```

```{r}
sum(c(1,2,3,NA))
sum(c(1,2,3,NA),na.rm = TRUE) # safe way
```

answer questions

-   what was your input ? how did you call the function?

-   what were you excepting? output,messages,other results?

-   what did you get?

-   how does what you get differ from what you were excepting ?

-   were you expectations correct in the first place ?

-   can you reproduce the problem (exactly)?

    ## Tools

    -   ***traceback*** : print out function call stack after an error
        occurs ,does nothing if there's no error

    -   ***debug : flags a function for debug mode which allow you to
        step through execution of a function se line at a time***

    -   browser : suspends execution of function wherever it is called
        and puts the funcion in debug mode

    -   trace : allows you to insert debugging code into a function a
        specific places

    -   recover : allows you to modify error behavior so that you can
        browse the function call stack
        
```{r}
    
    # mean(gx)
    # traceback()
    # 
    # lm(y~x)
    # traceback()
 ```





## simulation

rnorm(),dnorm(),pnorm(),rpois()

r : random number

d : density

p:cumulative distribution

q : quantile function

```{r}
rnorm(10) # mean =0 ,sd =1
rnorm(10,mean = 20, sd = 2) -> x# mean =0 ,sd =1
summary(x)
set.seed(1)
rnorm(5)

# qnorm(10,lower.tail = TRUE,log.p = FALSE)
# pnorm(10,lower.tail = TRUE,log.p = FALSE)
# dnorm(10,log = FALSE)
```

```{r}
rpois(10,5) -> x
summary(x)

ppois(2,2)  # prop[x <=2] 0.6766764
ppois(4,2)  # prop[x <=4] 0.947347


```

## Simulating a Linear Model

```{r}
set.seed(3)
x <- rnorm(100)
e <- rnorm(100)
y <- -0.5 + 2*x +e
plot(x,y)
```

## random sampling

```{r}
sample(1:10)
sample(1:10,6)

samples <- sample(3, size=1000, replace=TRUE, prob=c(.60,.20,.20)) # sampling data to groups 
train <- diamonds[which(samples == 1),c(1,5,6)] # filtering train 0.6 row ,selected cols
test <- diamonds[which(samples == 2), ]
valid <- diamonds[which(samples == 3), ]
head(train)
```

## profiling

-   how many time spend in different parts of programm

-   optimize code

-   profermance analyse or profiling

-   system.time()

elapsed time

paralle processing (threads)

```{r}
Rprof()
```

link your project to github using git 

