---
title: "Statistics"
author: "Bakro"
date: "12/27/2021"
output: 
  html_document:
   toc: true
   toc_float: true
   toc_depth: 3
   theme: flatly
   highlight: zenburn
   df_print: paged
   code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Probability

| Function | Purpose                      |
|----------|------------------------------|
| dnorm    | Normal density               |
| pnorm    | Normal distribution function |
| qnorm    | Normal quantile function     |
| rnorm    | Normal random variates       |

## distributions types

### Discrete distribution R

+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+
| name                            | rname  | params                                                                                                |
+=================================+========+=======================================================================================================+
| Binomial                        | binom  | n = number of trials; p = probability of success for one trial                                        |
+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+
| Geometric                       | geom   | p = probability of success for one trial                                                              |
+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+
| Hypergeometric                  | hyper  | m = number of white balls in urn; n = number of black balls in urn; k = numberof balls drawn from urn |
+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+
| Negative binomial (NegBinomial) | nbinom | size = number of successful trials; either prob = probability of successful trial                     |
|                                 |        |                                                                                                       |
|                                 |        | or mu = mean                                                                                          |
+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+
| Poisson                         | pois   | lambda = mean                                                                                         |
+---------------------------------+--------+-------------------------------------------------------------------------------------------------------+

### continuous distribution

+-------------------------+-------------+------------------------------------------------+
| Continuous distribution | R name      | Parameters                                     |
+=========================+=============+================================================+
| Beta                    | beta        | shape1; shape2                                 |
+-------------------------+-------------+------------------------------------------------+
| Cauchy                  | cauchy      | location; scale                                |
+-------------------------+-------------+------------------------------------------------+
| Chi-squared (Chisquare) | chisq       | df = degrees of freedom                        |
+-------------------------+-------------+------------------------------------------------+
| Exponential             | exp         | rate F f df1 and df2 = degrees of freedom      |
+-------------------------+-------------+------------------------------------------------+
| Gamma                   | gamma rate; | either rate or scale                           |
+-------------------------+-------------+------------------------------------------------+
| Log-normal (Lognormal)  | lnorm e     | meanlog = mean on logarithmic scale;           |
|                         |             |                                                |
|                         |             | sdlog = standard deviation on logarithmic scal |
+-------------------------+-------------+------------------------------------------------+
| Logistic                | logis       | location; scale                                |
+-------------------------+-------------+------------------------------------------------+
| Normal                  | norm        | mean; sd = standard deviation                  |
+-------------------------+-------------+------------------------------------------------+
| Student's t (TDist)     | t           | df = degrees of freedom                        |
+-------------------------+-------------+------------------------------------------------+
| Uniform                 | unif        | min = lower limit; max = upper limit           |
+-------------------------+-------------+------------------------------------------------+
| Weibull                 | weibull     | shape; scale                                   |
+-------------------------+-------------+------------------------------------------------+
| Wilcoxon                | wilcox      | m = number of observations in first sample;    |
|                         |             |                                                |
|                         |             | n = number of observations in second sample    |
+-------------------------+-------------+------------------------------------------------+

```{r}
library(ggplot2)
#?Normal
#?TDist
```

## Intro

### **Counting the Number of Combinations**

n!/r!(n − r)!

```{r}
choose(5,3) # How many ways can we select 3 items from 5 items?
# These numbers are also known as binomial coefficients.
```

### **Generating Combinations**

```{r}
combn(1:5,3)
```

```{r}
combn(c("T1","T2","T3","T4","T5"), 3)
```

### **Generating Random Numbers**

```{r}
runif(1)
rnorm(1)

```

```{r}
runif(1, min=-3, max=3) # One uniform variate between -3 and +3
```

```{r}
rnorm(1) # One standard Normal variate
```

```{r}
rnorm(10, mean=100, sd=15) # One Normal variate, mean 100 and SD 15
```

```{r}
rbinom(10, size=1, prob=0.5) # One binomial variate
```

```{r}
rpois(10, lambda=10) # One Poisson variate
```

```{r}
rexp(10, rate=0.1) # One exponential variate
```

```{r}
rgamma(10, shape=2, rate=0.1) # One gamma variate
```

```{r}
rnorm(3, mean=c(-10,0,+10), sd=1)
```

```{r}
means <- rnorm(100, mean=0, sd=0.2)
```

**Generating Reproducible Random Numbers**

```{r}
set.seed(666)
runif(10)
```

**Generating a Random Sample**

```{r}
n = seq(1990 , 2015 )
sample(n, 10)
```

**Generating Random Sequences**

sample(set, n, replace=TRUE)

```{r}
sample(c("H","T"), 10, replace=TRUE)
```

```{r}
sample(c(FALSE,TRUE), 20, replace=TRUE)
```

```{r}
sample(c(FALSE,TRUE), 20, replace=TRUE, prob=c(0.2,0.8))
```

```{r}
rbinom(10, 1, 0.8)
```

**Randomly Permuting a Vector**

sample(v, size=length(v), replace=FALSE)

```{r}
sample(1:10)
```

**Calculating Probabilities for Discrete Distributions**

**Problem**

You want to calculate either the simple or the cumulative probability associated with

a discrete random variable.

**Solution**

For a simple probability, P(X = x), use the density function. All built-in probability

distributions have a density function whose name is "d" prefixed to the distribution

name---for example, dbinom for the binomial distribution.

For a cumulative probability, P(X ≤ x), use the distribution function. All built-in probability

distributions have a distribution function whose name is "p" prefixed to the

distribution name; thus, pbinom is the distribution function for the binomial

distribution.

```{r}
# Suppose we have a binomial random variable X over 10 trials, where each trial has
#  a success probability of 1/2. Then we can calculate the probability of observing x = 7 by calling dbinom:

dbinom(7, size=10, prob=0.5)
```

```{r}
# The cumulative probability, P(X ≤ x), is given by the distribution function, 
# which is sometimes called the cumulative probability function. The distribution function for the binomial distribution is pbinom. Here is the cumulative probability for x = 7 (i.e., P(X ≤ 7)):
pbinom(7, size=10, prob=0.5)

# It appears the probability of observing X ≤ 7 is about 0.945.
```

```{r}
pbinom(7, size=10, prob=0.5, lower.tail=FALSE)
```

```{r}
#The interval probability, P(x1 < X ≤ x2), is the probability of observing X between the limits x1 and x2. It is simply calculated as the difference between two cumulative probabilities:
# P(X ≤ x2) − P(X ≤ x1). Here is P(3 < X ≤ 7) for our binomial variable:
pbinom(7,size=10,prob=0.5) - pbinom(3,size=10,prob=0.5)
```

```{r}
pbinom(c(3,7), size=10, prob=0.5)
```

```{r}
diff(pbinom(c(3,7), size=10, prob=0.5))
```

```{r}
binom.test(x=39, n=215, p=.15)


```

**Calculating Probabilities for Continuous Distributions**

**Problem**

You want to calculate the distribution function (DF) or cumulative distribution function

(CDF) for a continuous random variable.

**Solution**

Use the distribution function, which calculates P(X ≤ x). All built-in probability distributions

have a distribution function whose name is "p" prefixed to the distribution's

abbreviated name---for instance, pnorm for the Normal distribution.

**Discussion**

The R functions for probability distributions follow a consistent pattern, so the solution

to this recipe is essentially identical to the solution for discrete random variables

(Recipe 8.8). The significant difference is that continuous variables have no "probability"

at a single point, P(X = x). Instead, they have a density at a point.

Given that consistency, the discussion of distribution functions in Recipe 8.8 is

applicable here, too. The following table gives the distribution function for several

continuous distributions:

| Distribution     | Distribution function: P(X ≤ x) |
|------------------|---------------------------------|
| Normal           | pnorm(x, mean, sd)              |
| Student's t      | pt(x, df)                       |
| Exponential      | pexp(x, rate)                   |
| Gamma            | pgamma(x, shape, rate)          |
| Chi-squared (χ2) | pchisq(x, df)                   |

```{r}
# X ~ N(70, 3):
 pnorm(66, mean=70, sd=3)

```

```{r}
# Likewise, we can use pexp to calculate the probability that an exponential variable with a mean of 40 could be less than 20:
pexp(20, rate=1/40)
```

```{r}
# Just as for discrete probabilities, the functions for continuous probabilities 
# use lower.tail=FALSE to specify the survival function, P(X > x). This call to pexp gives the probability that the same exponential variable could be greater than 50:
pexp(50, rate=1/40, lower.tail=FALSE)
```

```{r}
# Also like discrete probabilities, the interval probability for a continuous variable, P(x1 <X < x2), is computed as the difference between two cumulative probabilities, P(X < x2) − P(X < x1). For the same exponential variable, here is P(20 < X < 50), the probability that it could fall between 20 and 50:
 
 pexp(50,rate=1/40) - pexp(20,rate=1/40)
```

**Converting Probabilities to Quantiles**

**Problem**

Given a probability p and a distribution, you want to determine the corresponding

quantile for p: the value x such that P(X ≤ x) = p.

**Solution**

Every built-in distribution includes a quantile function that converts probabilities to

quantiles. The function's name is "q" prefixed to the distribution name; thus, for instance,

qnorm is the quantile function for the Normal distribution.

The first argument of the quantile function is the probability. The remaining arguments

are the distribution's parameters, such as mean, shape, or rate:

```{r}
qnorm(0.05, mean=100, sd=15)
```

| Distribution     | Quantile function                                            |
|------------------|--------------------------------------------------------------|
| Binomial         | qbinom(p, size, prob)                                        |
| Geometric        | qgeom(p, prob)                                               |
| Poisson          | qpois(p, lambda)                                             |
| Normal           | qnorm(p, mean, sd)                                           |
| Student's t      | qt(p, df)                                                    |
| Exponential      | qexp(p, rate)                                                |
| Gamma            | qgamma(p, shape, rate=rate) or qgamma(p, shape, scale=scale) |
| Chi-squared (χ2) | qchisq(p, df)                                                |

**Plotting a Density Function**

```{r}
x <- seq(from=0, to=6, length.out=100) # Define the density domains
ylim <- c(0, 0.6)
par(mfrow=c(2,2)) # Create a 2x2 plotting area
plot(x, dunif(x,min=2,max=4), main="Uniform", # Plot a uniform density
type='l', ylim=ylim)
plot(x, dnorm(x,mean=3,sd=1), main="Normal", # Plot a Normal density
type='l', ylim=ylim)
plot(x, dexp(x,rate=1/2), main="Exponential", # Plot an exponential density
type='l', ylim=ylim)
plot(x, dgamma(x,shape=2,rate=1), main="Gamma", # Plot a gamma density
type='l', ylim=ylim)

```

```{r}
x <- seq(from=-3, to=+3, length.out=100)
y <- dnorm(x)
plot(x, y, main="Standard Normal Distribution", type='l',
ylab="Density", xlab="Quantile")
abline(h=0)

# The body of the polygon follows the density curve where 1 <= z <= 2
region.x <- x[1 <= x & x <= 2]
region.y <- y[1 <= x & x <= 2]
# We add initial and final segments, which drop down to the Y axis
region.x <- c(region.x[1], region.x, tail(region.x,1))
region.y <- c( 0, region.y, 0)

polygon(region.x, region.y, density=-1, col="red")
```

```{r}

```

## Appendix

```{r}
x = seq(from = -10,to = 10 ,by = 0.1)
y = dnorm(x)
df = data.frame(x ,y)
head(df)
ggplot(data = df ,mapping =  aes(x = x  )) + 
  geom_density()
```

```{r}
dnorm(0) == 1/sqrt(2*pi)
dnorm(1) == exp(-1/2)/sqrt(2*pi)
dnorm(1) == 1/sqrt(2*pi*exp(1))

## Using "log = TRUE" for an extended range :
par(mfrow = c(2,1))
plot(function(x) dnorm(x, log = TRUE), -60, 50,
     main = "log { Normal density }")
curve(log(dnorm(x)), add = TRUE, col = "red", lwd = 2)
mtext("dnorm(x, log=TRUE)", adj = 0)
mtext("log(dnorm(x))", col = "red", adj = 1)

plot(function(x) pnorm(x, log.p = TRUE), -50, 10,
     main = "log { Normal Cumulative }")
curve(log(pnorm(x)), add = TRUE, col = "red", lwd = 2)
mtext("pnorm(x, log=TRUE)", adj = 0)
mtext("log(pnorm(x))", col = "red", adj = 1)

## if you want the so-called 'error function'
erf <- function(x) 2 * pnorm(x * sqrt(2)) - 1
## (see Abramowitz and Stegun 29.2.29)
## and the so-called 'complementary error function'
erfc <- function(x) 2 * pnorm(x * sqrt(2), lower = FALSE)
## and the inverses
erfinv <- function (x) qnorm((1 + x)/2)/sqrt(2)
erfcinv <- function (x) qnorm(x/2, lower = FALSE)/sqrt(2)
```

```{r}
par(mfrow=c(3,2)) # Create a 2x2 plotting area
x1 =rexp(100 , rate = 0.1)
plot(density(x1))
x = rgamma(n = 100 , shape = 2 , rate = 0.1)
plot(density(x))
x1 =runif(100 )
plot(density(x1))
x1 = rnorm(100)
plot(density(x1))


x1 = rlnorm(10000) # log norm 
plot(density(x1))

rnbinom(n =1000 , size = 2 , prob  = .5)

x1 = rpois(100 , lambda = 1)
plot(density(x1))

```

```{r}
x <- seq(from=0, to=6, length.out=100) # Define the density domains
ylim <- c(0, 0.6)
par(mfrow=c(2,2)) # Create a 2x2 plotting area
plot(x, dunif(x,min=2,max=4), main="Uniform", # Plot a uniform density
type='l', ylim=ylim)
plot(x, dnorm(x,mean=3,sd=1), main="Normal", # Plot a Normal density
type='l', ylim=ylim)
plot(x, dexp(x,rate=1/2), main="Exponential", # Plot an exponential density
type='l', ylim=ylim)

z = seq(from=-10, to=10, by =1)
plot(z , dpois(z,lambda = 1), main="Gamma", # Plot a gamma density
type='l', ylim=ylim)




```
