---
title: "List"
author: "Bakro"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Purr

## Map Functions

### ONE LIST

**map(.x, .f, ...) Apply a function to each element**

**of a list or vector, return a list.**

```{r}
x <- list(1:10, 11:20, 21:30)
l1 <- list(x = c("a", "b"), y = c("c", "d"))
map(l1, sort, decreasing = TRUE)
```

#### map_dbl(.x, .f, ...)

***Return a double vector.***

```{r}
map_dbl(x, mean)
```

#### map_int(.x, .f, ...)

***Return an integer vector.***

```{r}
map_int(x, length)
```

#### map_chr(.x, .f, ...)

***Return a character vector.***

```{r}
map_chr(l1, paste, collapse = "")
```

#### map_lgl(.x, .f, ...)

***Return a logical vector.***

```{r}
map_lgl(x, is.integer)
```

#### map_dfc(.x, .f, ...)

***Return a data frame created by column-binding.***

```{r}
map_dfc(l1, rep, 3)
```

#### map_dfr(.x, .f, ..., .id = NULL)

***Return a data frame created by row-binding.***

```{r}
map_dfr(x, summary)
```

#### walk(.x, .f, ...)

***Trigger side effects, return invisibly.***

```{r}
walk(x, print)
```

### TWO LISTS

***map2(.x, .y, .f, ...) Apply a function to pairs of***

***elements from two lists or vectors, return a list.***

```{r}
y <- list(1, 2, 3); z <- list(4, 5, 6); l2 <- list(x = "a", y = "z")

map2(x, y, ~ .x * .y) # element wise
```

#### map2_dbl(.x, .y, .f, ...)

***Return a double vector.***

```{r}
map2_dbl(y, z, ~ .x / .y) # element wise
```

#### map2_int(.x, .y, .f, ...)

***Return an integer vector.***

```{r}
list1 <- list(1L,2L) ;list2 <- list(3L,4L)
map2_int(list1, list2,~ .x + .y)  # error
```

#### map2_chr(.x, .y, .f, ...)

***Return a character vector.***

```{r}
map2_chr(l1, l2, paste,collapse = ",", sep = ":")
```

#### map2_lgl(.x, .y, .f, ...)

***Return a logical vector.***

```{r}

map2_lgl(l2, l1, `%in%`)
```

***map2_dfc(.x, .y, .f, ...)***

***Return a data frame created by column-binding.***

```{r}
map2_dfc(l1, l2,~ as.data.frame(c(.x, .y)))
```

#### map2_dfr(.x, .y, .f, ..., .id =NULL)

Return a data frame created by row-binding.

```{r}
map2_dfr(l1, l2,

~ as.data.frame(c(.x, .y)))
```

#### walk2(.x, .y, .f, ...) Trigger

side effects, return invisibly.

```{r}
#walk2(objs, paths, save)
```

### MANY LISTS

pmap(.l, .f, ...) Apply a function to groups of

elements from a list of lists or vectors, return a list.

```{r}
pmap(list(x, y, z), ~ ..1 * (..2 + ..3))
```

#### pmap_dbl(.l, .f, ...)

***Return a double vector.***

```{r}
pmap_dbl(list(y, z), ~ .x / .y)
```

#### pmap_int(.l, .f, ...)

***Return an integer vector.***

```{r}
pmap_int(list(list1, list2), `+`)    # error
```

#### pmap_chr(.l, .f, ...)

***Return a character vector.***

```{r}
pmap_chr(list(l1, l2), paste,collapse = ",", sep = ":")
```

#### pmap_lgl(.l, .f, ...)

***Return a logical vector.***

```{r}
pmap_lgl(list(l2, l1), `%in%`)
```

#### pmap_dfc(.l, .f, ...)

***Return a data frame created bycolumn-binding.***

```{r}
pmap_dfc(list(l1, l2),~ as.data.frame(c(.x, .y)))
```

#### pmap_dfr(.l, .f, ..., .id =NULL)

***Return a data frame created by row-binding.***

```{r}
pmap_dfr(list(l1, l2),~ as.data.frame(c(.x, .y)))
```

#### pwalk(.l, .f, ...)

***Trigger side effects, return invisibly.***

```{r}
#pwalk(list(objs, paths), save)
```

### LISTS AND INDEXES

imap(.x, .f, ...)

Apply .f to each element and its index, return a list.

```{r}
imap(y, ~ paste0(.y, ": ", .x))
```

## Work with Lists

### Filter

#### keep(.x, .p, ...)

Select elements that pass a logical test.

Conversely, discard().

```{r}
temp <- list(NA,3,NA)
keep(temp, is.na)
```

### compact(.x, .p = identity)

Drop empty elements.

```{r}
temp <- list(NULL,3,NA)
compact(temp)
```

### head_while(.x, .p, ...)

Return head elements until one does not pass.

Also tail_while().

```{r}
temp <- list("3",1,"4")
head_while(temp, is.character)
```

### detect

detect(.x, .f, ..., dir =c("forward", "backward"),

.right = NULL, .default = NULL)

Find first element to pass.

```{r}
detect(temp, is.character)
```

## detect_index

detect_index(.x, .f, ..., dir =c("forward", "backward"),.right = NULL)

Find index of first element to pass.

```{r}
 detect_index(temp, is.character)
```

### every(.x, .p, ...)

Do all elements pass a test?

```{r}
every(temp, is.character)
temp2 <- list("c")
every(temp2, is.character)


```

### some(.x, .p, ...)

Do some elements pass a test?

```{r}
some(temp, is.character)
```

### none(.x, .p, ...)

Do no elements pass a test?

```{r}
none(temp, is.character)
temp2 <-  list(1L)
none(temp2, is.character)
```

### has_element(.x, .y)

Does a list contain an element?

```{r}
has_element(temp, "3")
```

### vec_depth(x)

Return depth (number of levels

of indexes).

```{r}
vec_depth(temp)
temp
```

### Index

#### pluck(.x, ..., .default=NULL)

Select an element by name or index. Also attr_getter() and chuck().

```{r}
temp <- list(first ="1",second = 2L ,TRUE ,list(1,2))
pluck(temp, 1)
temp %>% pluck(first)
pluck(temp,first)
temp$firstelement
temp[[1]]
```

### assign_in(x, where, value)

Assign a value to a location using pluck selection.

```{r}
assign_in(temp, "first", 5)

# temp  %>% assign_in("first", 5)
```

#### modify_in(.x, .where, .f)

Apply a function to a value at

a selected location.

```{r}
modify_in(temp, "second", abs)

# temp %>% modify_in("second", abs)
```

### Reshape

#### flatten(.x) 

Remove a level of indexes from a list.

Also flatten_chr() etc.

```{r}
flatten(temp)
```

array_tree(array, margin = NULL) Turn array into list.

Also array_branch().

```{r}
array_tree(matrix(1:9,nrow = 3), margin = 2) # 1 row 2 col ,....
```

#### cross2(.x, .y, .filter = NULL)

All combinations of .x and .y.

Also cross(), cross3(), and cross_df().

```{r}
cross2(1:2, 4)
```

#### transpose(.l, .names = NULL)

Transposes the index order in

a multi-level list.

```{r}
temp2 <- list(1:2,3:4,5:6)
transpose(temp2)
```

#### set_names(x, nm = x)

Set the names of a vector/list

directly or with a function.

```{r}
set_names(temp2, c("p", "q", "r"))
set_names(temp2, tolower)
```

### Modify

#### modify(.x, .f, ...) 

Apply afunction to each element. Also

modify2(), and imodify().

```{r}
modify(x, ~.+ 2)
```

#### modify_at(.x, .at, .f, ...)

Apply a function to selected elements.

Also map_at().

modify_at(x, "b", \~.+ 2)

#### modify_if(.x, .p, .f, ...) 

Apply a function to elements that pass

a test. Also map_if().

```{r}
modify_if(x, is.numeric,~.+2)
```

#### modify_depth(.x, .depth, .f, ...)

Apply function to each element

at a given level of a list. Alsomap_depth().

```{r}
modify_depth(y, 2, ~.+ 2)
```

### Combine

```{r}
# Add values to end of list.
# append(list1, list2, after =length(list1))
append(list1, list(d = 1))
# prepend(x, values, before = 1)
# Add values to start of list.
prepend(list1, list(d = 1))
# splice(…) Combine objects into
# a list, storing S3 objects as sublists.
splice(list1, list2, "foo")
```

### List-Columns

List-columns are columns of a data frame where each element is a list or vector instead of an atomic

value. Columns can also be lists of data frames. See tidyr for more about nested data and list columns.

WORK WITH LIST-COLUMNS

Manipulate list-columns like any other kind of column, using dplyr functions like mutate() and

transmute(). Because each element is a list, use map functions within a column function to

manipulate each element.

map(), map2(), or pmap() return lists and will

create new list-columns.

```{r}
# starwars %>%
# transmute(ships = map2(vehicles,
# starships, append)
```

Suffixed map functions like map_int() return an atomic data type and

will simplify list-columns into regular columns.

```{r}
starwars %>%
mutate(n_films = map_int(films, length))
```

### Reduce

reduce(.x, .f, ..., .init, .dir = c("forward",

"backward")) Apply function recursively to each

element of a list or vector. Also reduce2().

```{r}
reduce(x, sum)
```

accumulate(.x, .f, ..., .init) Reduce a list, but also

return intermediate results. Also accumulate2().

```{r}
accumulate(x, sum)
```
