---
title: "Factors"
author: "Abdulmajid Bakro"
date: "6/10/2021"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r   include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# klippy::klippy()
```

# Factors 
## forcats

the **forcats** package provides tools for working with factors ,which
are R's data structure for categorial data (datatype for factors always
numeric ) <br>

```{r  message=FALSE}
# klippy::klippy(position = c('center', 'right'))
library(forcats)
library(ggplot2)
library(dplyr)
```

factors : R represent categorical data with factors . A factor is an
integer vector with a **levels** attribute that stores a set of mappings
between integers and categorical values ,when you view a factor , R
display not the integers , burt the levels associated with them <br>
factor(x=character(),levels,labels = levels, exclude = NA , ordered =
is.ordered(x),nmax = NA)

```{r create a factor with factor()}

f <- factor(c('a','b','c','c','b','a'),levels = c('a','b','c') )
```

to convert a vector to a factor using as_factor()

```{r convert factor }

vec <- c('a','b','c','c','b','a')
fac <-as_factor(vec)
fac
```

return levels

```{r}
levels(fac)
```

structure of factors

```{r}
 summary(fac)
unclass(fac)
```

# inspect Factors

count number of values with each levels

```{r}
fct_count(fac,sort =FALSE ,prop = FALSE)
fct_count(iris$Species,sort =FALSE ,prop = FALSE)
```

check for levels in factor

```{r}
fct_match(fac,"a")
```

return the unique values , removing duplicate

```{r}
fct_unique(fac)
```

## combine Factors 

fct_c(...) combine factors with different levels Also **fct_cross**

```{r}
fct1 <- factor(c("a","b"))
fct2 <- factor(c("c","b"))
fct_c(fct1,fct2)
```

fct_unify(fs, levels =lvls_union(fs)) Standardizelevels across a list of
factors. <br>

```{r}
fs <- list(factor("a"), factor("b"), factor(c("a", "b")))
fct_unify(fs)
```

## change order of levels {#change_order_levels}

```{r intro to order factors}
faInt <- factor(c(0L, 1L, 2L, 2L, 0L, 1L))
faIntOrd <- ordered(faInt)
faIntOrd
faIntOrd < 2 # only given values
faIntOrd < 3 # NAs
```

fct_relevel(.f, ..., after = 0L) Manually reorder factor levels.

```{r}
fct_relevel(f, c("b", "c", "a"))
fct_relevel(f, "a", after = 1)
fct_relevel(f, "a", after = Inf)
```

fct_infreq(f, ordered = NA) <br> Reorder levels by the frequency in
which they appear in the data (highest frequency first). Also
fct_inseq().

```{r}
f3 <- factor(c("c", "c", "a"))
fct_infreq(f3)

```

fct_inorder(f, ordered = NA) Reorder levels by order in which they
appear in the data.

```{r}
f2 <- factor(c("b","a","c"))
fct_inorder(f2)
```

fct_rev(f) Reverse level order.

```{r}
f4 <- factor(c("a","b","c"))
fct_rev(f4)
```

fct_shift(f) Shift levels to left or right, wrapping around end.

```{r}
f4 <- factor(c("a","b","c"))
fct_shift(f4)
```

fct_shuffle(f, n = 1L) Randomly permute order of factor levels.
fct_shuffle(f4)

```{r}
f4 <- factor(c("a","b","c"))
fct_shuffle(f4)
```

fct_reorder(.f, .x, .fun = median, ..., .desc = FALSE) Reorder levels by
their relationship with another variable.

```{r}
head(PlantGrowth)
boxplot(PlantGrowth$weight ~ PlantGrowth$group)

boxplot(data = PlantGrowth,
weight ~ reorder(group, weight))
```

```{r}
faGeschl <- factor(c(rep("W", each = 10), 
                     rep("M", each = 10), 
                     rep("D", each = 10)))
faGeschl
vnGew_W <- c(58.10, 54.65, 58.03, 56.83, 60.44, 65.92, 61.38, 59.10, 
             59.66, 54.65)
vnGew_M <- c(84.84, 80.02, 78.80, 82.45, 81.30, 77.47, 82.61, 85.35, 
             78.41, 81.62)
vnGew_D <- c(67.53, 67.31, 72.68, 69.65, 63.38, 2.14, 63.13, 71.31, 
             75.00, 69.72)
vnGew <- c(vnGew_W, vnGew_M, vnGew_D)

faGeschlReOrd <- reorder(faGeschl, vnGew, FUN = mean)
faGeschlReOrd
```

fct_reorder2(.f, .x, .y, .fun = last2, ..., .desc = TRUE) Reorder levels
by their final values when plotted with two other variables.

```{r}
ggplot(diamonds,aes(carat, price,
color = fct_reorder2(color, carat,
price))) + geom_smooth(ls = F)
```

## Change the value of levels

fct_recode(.f, ...) Manually change levels. Also fct_relabel() which
obeys purrr::map syntax to apply a function or expression to each level.

```{r}
f <- factor(c("a","b","c"))
fct_recode(f, v = "a", x = "b", z = "c")
fct_relabel(f, ~ paste0("x", .x))
```

fct_anon(f, prefix = "") Anonymize levels with random integers.

```{r}
fct_anon(f)
```

fct_collapse(.f, ..., other_level = NULL) Collapse levels into manually
defined groups.

```{r}
fct_collapse(f, x = c("a"))
```

fct_lump_min(f, min, w = NULL, other_level = "Other") Lumps together
factors that appear fewer than min times. Also fct_lump_n(),
fct_lump_prop(), and fct_lump_lowfreq(). fct_lump_min(f, min = 2)

```{r}
f2 <- factor(c("a","a","a","b","c"))
fct_lump_min(f2, min = 2)
```

fct_other(f, keep, drop, other_level = "Other") Replace levels with
"other."

```{r}
fct_other(f, keep = c("a", "b"))
```

## add or drop levels 

fct_drop(f, only) Drop unused levels.

```{r}
f5 <- factor(c("a","b"),c("a","b","x"))
f6 <- fct_drop(f5)
f5 ; f6
```

fct_expand(f, ...) Add levels to a factor.

```{r}
fct_expand(f6, "x")
```

fct_explicit_na(f, na_level="(Missing)") Assigns a level to NAs to
ensure they appear in plots, etc.

```{r}

fct_explicit_na(factor(c("a", "b", NA)))
```

## operations 

sampling or indexing

```{r}
fac[2:4]
fac[c(1, 3, 5)]
fac[-c(2,4)]
```

sampling using operation

```{r}
fac < 1   #  ‘<’ not meaningful for factors
```

add values to factors

```{r}
fac <- factor(c("a","b","c"))
fac[4] <- "a"
fac
fac[5] <- "Z"  # z not found in levels 
fac <-fct_expand(fac,"Z")
fac[5] <- "Z"
fac
```

```{r}
faLH <- gl(n = 2, k = 3, labels = c("low", "high"))
faLH
```

grid or cross factors

```{r}
fac1 <- factor(c(1,2,3))
fac2 <- factor(c(1,2,3))
dfFactKombi <- expand.grid(fac1, fac2)
dfFactKombi
intersected <- interaction(fac1, fac2)
intersected
```

```{r eval=TRUE}
fact1 <- factor(rep(c(1:10),3))

```

## Examples

###     Months orders  

```{r}
x2 <- c("Dec", "Apr", "Jam", "Mar")
```

```{r}
sort(x2)

```

```{r}
month_levels <- c(
  "Jan", "Feb", "Mar", "Apr", "May", "Jun", 
  "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"
)
```

```{r}
y1 <- factor(x2, levels = month_levels)
y1
#> [1] Dec Apr Jan Mar
#> Levels: Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec
sort(y1)
#> [1] Jan Mar Apr Dec
#> Levels: Jan Feb Mar Apr May Jun Jul Aug Sep Oct Nov Dec
```

### Ranking

```{r}
relig_summary <- gss_cat %>%
  group_by(relig) %>%
  summarise(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )


ggplot(relig_summary, aes(tvhours, relig)) + geom_point()
```

```{r}
ggplot(relig_summary, aes(tvhours, fct_reorder(relig, tvhours))) +
  geom_point()
```

## RECAP

### 1-what is factors

factors: R represent categorical data with factors . A factor is an
integer vector with a **levels** attribute that stores a set of mappings
between integers and categorical values ,when you view a factor , R
display not the integers , burt the levels associated with them

### 2- initialize  factors

```{r}
fac1 <-factor(c("SWE", "UK", "SWE", "IRE", "SWE", "THA", "IRE", "IND", "KOR", "IND"))
fac2 <-factor(c("BRA", "VEN", "MEX", "BRA", "MEX", "ARG", "CHI", "ARG", "MEX"))
fac1
fac2
 
```

```{r}
fac3 <-factor(c("BRA", "VEN", "MEX", "BRA", "MEX", "ARG", "CHI", "ARG", "MEX"),
              levels = c("ARG" , "BRA" , "CHI", "MEX", "VEN"))
fac3

fac4 <-factor(c("BRA", "VEN", "MEX", "BRA", "MEX", "ARG", "CHI", "ARG", "MEX"),
              levels = c("ARG" , "BRA" , "CHI"))
fac4
```

::: {style="color:green ;"}
when you don't give levels --> distinct levels ordered alphabet

when we give all levels no problem

when we give part of levels replace element that don't have levels with
NA
:::

### 3- combine factors 

```{r}
combinfactor <- fct_c(fac1,fac2)
fct_unify(list(fac1,fac2))
fct_unify(list(fac1,fac2),levels = lvls_union(list(fac1,fac2)))
```

### 4-  levels & count of levels

```{r}
levels(combinfactor)
nlevels(combinfactor)
```

### 5- indexing 

```{r}
fac1[1] # first element 
fac2[1]
fac1[length(fac1)]  # last element 
fac2[length(fac2)]
```

### 6- inserting

you can only insert values that found in levels (constrains),when you
want to add element that

don't have levels **first add level** than **add element**

```{r}
fac1[length(fac1)+1] <- "KOL"

```

```{r}
fac1 <-fct_expand(fac1,"KOL")
fac1[length(fac1)+1] <- "KOL"
fac1
```

### 7- Change values of levels 

```{r}
levels(fac1)
fac3 <-fct_relabel(fac1,~paste0(tolower(levels(fac1))))
fac3
# fac4 <-fct_recode(fac1,tolower(levels(fac1)))
```

### 8- drop levels

you should remove all observation or records with these levels

```{r}
fac2
fac2[fac2 == "BRA" ||fac2 == "MEX"] <- NA
fct_drop(fac2)
#fac3 <- fct_drop(fac2,c("BRA","MEX"))
#fac3

```

### 9- order levels 

```{r}
fac1 <- factor(c("c1","c3","c2","c4","IRE"))
vec <- c(1.3,1.5,1.4,1.8,1.45)
#fac3 <- reorder(fac1 ,vec~fac1)
fac3 <- fct_reorder(.f = fac1 , vec ,min)
fac3
 
```

```{r}
boxplot(Sepal.Width ~ Species, data = iris)
boxplot(Sepal.Width ~ fct_reorder(Species, Sepal.Width, .desc = TRUE), data = iris)
```

### 10-  compare operation with ordered factors 

```{r}

fac3 <- ordered(fac3)
ind <- fac3 < "IRE";
fac3[ind]
```



