---
title: "Tidyr"
author: "Bakro"
date: "10/5/2021"
output: 
  html_document:
    toc: true
    toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r  setup,klippy, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo = FALSE , message=FALSE ,warning=FALSE}
library(tidyverse)
```

# Basics

## Tidy data

**Tidy data** is a way to organize tabular data in a

consistent data structure across packages.

A table is tidy if:

-   Each variable is in its own column

-   Each observation, or case, is in its own row

-   Access variables as vectors

-   Preserve cases in vectorized operations

## **Tibbles**

Tibbles are a table format provided

by the tibble package. They inherit the

data frame class, but have improved behaviors:

• Subset a new tibble with ], a vector with [[ and \$.

• No partial matching when subsetting columns.

• Display concise views of the data on one screen.

options(tibble.print_max = n, tibble.print_min = m,

tibble.width = Inf) Control default display settings.

View() or glimpse() View the entire data set.

## CONSTRUCT A TIBBLE

**tibble(...)** Construct by columns.

```{r}
tibble(x = 1:3, y = c("a", "b", "c"))
```

**tribble(...)** Construct by rows.

```{r}
tribble(~x, ~y,

1, "a",

2, "b",

3, "c")
```

**as_tibble(x, ...)**

Convert a data frame to a tibble.

```{r}
as_tibble(diamonds)
is_tibble(diamonds)
```

**enframe**(x, name = "name", value = "value")

Convert a named vector to a tibble. Also deframe().

**is_tibble(x)** Test whether x is a tibble.

# Reshape Data

> Pivot data to reorganize values into a new layout.

## pivot_longer

pivot_longer(data, cols, names_to = "name",values_to = "value",
values_drop_na = FALSE)

"Lengthen" data by collapsing several columnsinto two. Column names move
to a new

names_to column and values to a new values_to column.

```{r}
head(table4a)
pivot_longer(table4a, cols = 2:3, names_to ="year",

values_to = "cases")
```

## pivot_wider()

pivot_wider(data, names_from = "name",

values_from = "value")

The inverse of pivot_longer(). "Widen" data by

expanding two columns into several. One column

provides the new column names, the other the

values.

```{r}
head(table2)
pivot_wider(table2, names_from = type,

values_from = count)
```

# Split Cells

> Use these functions to split or combine cells into individual,
> isolated values.

## unite()

unite(data, col, ..., sep = "\_", remove = TRUE,

na.rm = FALSE) Collapse cells across several

columns into a single column.

```{r}
head(table5)
unite(table5, century, year, col = "year", sep = "")
```

## separate()

separate(data, col, into, sep = "$$\^\[:alnum:$$]+",

remove = TRUE, convert = FALSE, extra = "warn",

fill = "warn", ...) Separate each cell in a column

into several columns. Also **extract().**

```{r}
head(table3)
separate(table3, rate, sep = "/",

into = c("cases", "pop"))
```

## separate_rows()

separate_rows(data, ..., sep = "",convert = FALSE) Separate each cell in
a column into several rows.

```{r}
head(table3)
separate_rows(table3, rate, sep = "/")
```

# Expand Tables

> Create new combinations of variables or identify implicit missing
> values (combinations of
>
> variables not present in the data).

## expand(data, ...)

Create a new tibble with all possible combinations of the values

of the variables listed in ...

Drop other variables.

```{r}
as_tibble(mtcars)
expand(mtcars, cyl, gear,carb)
```

## complete()

complete(data, ..., fill =list()) Add missing possible

combinations of values of variables listed in ... Fill

remaining variables with NA.

```{r}
complete(mtcars, cyl, gear,carb)
```

# Handle Missing Values

> Drop or replace explicit missing values (NA).

## drop_na()

drop_na(data, ...) Drop rows containing NA's in ...columns.

```{r}
x1<- c("a","b","c","d","e")
x2<- c(1,NA,NA,3,NA)
x <- data.frame(x1,x2)
drop_na(x, x2)
```

## fill() 

fill(data, ..., .direction ="down") Fill in NA's in ...

columns using the next or previous value.

```{r}
fill(x, x2)
```

## replace_na()

replace_na(data, replace)

Specify a value to replace

NA in selected columns.

```{r}
replace_na(x, list(x2 = 2))
```

# Nested Data

A nested data frame stores individual tables as a list-column of data
frames within a larger organizing data frame. List-columns can also be
lists of vectors or lists of varying data types.

Use a nested data frame to:

• Preserve relationships between observations and subsets of data.
Preserve the type of the variables being nested (factors and datetimes
aren't coerced to character).

• Manipulate many sub-tables at once with purrr funcitons like map(),
map2(), or pmap() or with dplyr rowwise() grouping.

## CREATE NESTED DATA

nest(data, ...) Moves groups of cells into a list-column of a data

frame. Use alone or with **dplyr::group_by()**:

1\. Group the data frame with group_by() and use nest() to move

the groups into a list-column.

```{r}
as_tibble(storms)
n_storms <- storms %>%

group_by(name) %>%

nest()
as_tibble(n_storms)
```

2\. Use nest(new_col = c(x, y)) to specify the columns to group

using dplyr::select() syntax.

```{r}
n_storms <- storms %>%

nest(data = c(year:long))
as_tibble(n_storms)
```

```{r}
#Index list-columns with [[]]. 
n_storms$data[[1]]
```

## CREATE TIBBLES WITH LIST-COLUMNS

```{r}
#tibble::tribble(…) Makes list-columns when needed.
tribble( ~max, ~seq,
3, 1:3,
4, 1:4,
5, 1:5)
```

**tibble::tibble(...)** Saves list input as list-columns.

```{r}
tibble(max = c(3, 4, 5), seq = list(1:3, 1:4, 1:5))
```

**tibble::enframe**(x, name="name", value="value")

Converts multi-level list to a tibble with list-cols.

```{r}
enframe(list('3'=1:3, '4'=1:4, '5'=1:5), 'max', 'seq')
```

## OUTPUT LIST-COLUMNS FROM OTHER FUNCTIONS

dplyr::mutate(), transmute(), and summarise() will output

list-columns if they return a list.

```{r}
mtcars %>%

group_by(cyl) %>%

summarise(q = list(quantile(mpg)))
```

## RESHAPE NESTED DATA

### unnest()

unnest(data, cols, ..., keep_empty = FALSE) Flatten nested columns

back to regular columns. The inverse of nest().

```{r}
as_tibble(n_storms %>% unnest(data))
```

### unnest_longer()

unnest_longer(data, col, values_to = NULL, indices_to = NULL)

Turn each element of a list-column into a row.

```{r}
starwars %>%
select(name, films) %>%
unnest_longer(films)
```

### unnest_wider()

unnest_wider(data, col) Turn each element of a list-column into a

regular column.

```{r  eval = FALSE}
starwars %>%
select(name, films) %>%
unnest_wider(films)
```

### hoist()

hoist(.data, .col, ..., .remove = TRUE) Selectively pull list components

out into their own top-level columns. Uses purrr::pluck() syntax for

selecting from lists.

```{r}
as_tibble(starwars)
starwars %>%
select(name, films) %>%
hoist(films, first_film = 1, second_film = 2)
```

## TRANSFORM NESTED DATA

A vectorized function takes a vector, transforms each element in

parallel, and returns a vector of the same length. By themselves

vectorized functions cannot work with lists, such as list-columns.

**dplyr::rowwise**(.data, ...) Group data so that each row is one

group, and within the groups, elements of list-columns appear

directly (accessed with [[ ), not as lists of length one. When you

use rowwise(), dplyr functions will seem to apply functions to

list-columns in a vectorized fashion.

-   Apply a function to a list-column and **create a new list-column.**

```{r}
as_tibble(n_storms)
n_storms %>%
rowwise() %>%
mutate(n = list(dim(data)))
```

-   Apply a function to a list-column and **create a regular column.**

```{r}
as_tibble(n_storms)
n_storms %>%
rowwise() %>%
mutate(n = nrow(data))
```

Collapse **multiple list-columns** into a single list-column.

```{r}
as_tibble(n_storms)
starwars %>%
rowwise() %>%
mutate(transport = list(append(vehicles, starships)))
```

Apply a function to **multiple list-columns.**

```{r}
as_tibble(n_storms)
starwars %>%
rowwise() %>%
mutate(n_transports = length(c(vehicles, starships)))
```

***See purrr package for more list functions.***

```{r}

# diamonds2 <- rep(diamonds$cut,100000)
# 
# class(diamonds2)
# object.size(diamonds2)
# object.size(as.vector(diamonds2))
# 
# diamonds2 <- as.data.frame(diamonds2)
# colnames(diamonds2)
# start <- proc.time()
# final_result <- filter(diamonds2 ,diamonds2 =="Good")
# end <- proc.time()
# 
# print(end - start)
# 
# diamonds2$diamonds2 <- as.vector(diamonds2$diamonds2)
# start <- proc.time()
# final_result <- filter(diamonds2 ,diamonds2 =="Good")
# end <- proc.time()
# 
# print(end - start)
```

# Example

## example1

```{r}
load(url("http://dwoll.de/rKompakt/dat.Rdata"))
# have avg of attention ,verbal ,DV
dataavg1 <- datL %>% group_by(ID) %>% summarise(
  attention = mean(attention,na.rm = TRUE) ,
  verbal = mean(verbal,na.rm = TRUE),
  DV = mean(DV,na.rm = TRUE)
) %>% ungroup()
# from long to wide table depend on variables 
data2 <- pivot_wider(datL, names_from = c(phase), 
                     values_from = c(attention,verbal,DV) )

# need to avg of DV_pre ,DV_post ,DV_fup
dataavg2 <- data2 %>% group_by(ID) %>% summarise(
  DV_pre = mean(DV_pre,na.rm = TRUE) ,
  DV_post = mean(DV_post,na.rm = TRUE),
  DV_fup = mean(DV_fup,na.rm = TRUE)
) %>% ungroup()

data3 <- pivot_wider(data2, names_from = c(hour), 
                     
                     values_from = c(DV_pre ,DV_post ,DV_fup ) )

# join dataavg1,data3,dataavg2

rm(data2)
t <- cbind(dataavg1,dataavg2,data3)
t <- t[,-c(5,9)] # repeated id col



x <- c(1,3, 5) 
y <- c(3, 2, 10)
cbind(x,y)
```

## example2

### import data

```{r}
vcFile <- 
"https://raw.githubusercontent.com/excelsiordata/DATA607/master/Discussion5.csv" 

dfShipping <- 
   read.csv(vcFile, na.strings=c("","NA"), stringsAsFactors = FALSE)
```

save copy of raw data

```{r}
dfshipping_copy <- dfShipping
```

### tidy data

here monitoring colnames and datatype

```{r}
str(dfshipping_copy)
```

```{r}
# clean types of variables 
 dfshipping_copy$China <- str_replace(dfshipping_copy$China, "#N/A", "") 
 dfshipping_copy$China <- as.numeric(dfshipping_copy$China) 
 dfshipping_copy <- tidyr::fill(dfshipping_copy, Date) 
 dfshipping_copy$Date <- as.Date(dfshipping_copy$Date, tryFormats = c("%m/%d/%y", "%Y-%m-%d")) 
```

reshaping data

```{r}
dfshipping_copy <- tidyr::pivot_longer(dfshipping_copy, cols=3:ncol(dfshipping_copy), names_to = "country", values_to = "price")
dfshipping_copy <- tidyr::pivot_wider(dfshipping_copy, names_from = "Values", values_from = "price") 
```



rename columns and relocated

```{r}
colnames(dfshipping_copy) <- c("date","country","price_of_carrier","shipping_fees_collected")
  # relocate columns
  dfshipping_copy <- dfshipping_copy %>%
    relocate(country)
```

# rename Values in columns

```{r}
dfshipping_copy <- dplyr::mutate(dfshipping_copy, country=case_when(
   country=="Tawian"~"Taiwan",
   country=="US"~"USA",
   country=="Singapore"~"Singapur",
   country=="Canada"~"Kanada",
   country=="Hong.Kong"~"Hongkong",
   country=="England"~"England",
   country=="China"~"China",
   country== "Korea"~"Korea"))
```

```{r}
tishipping_sort <- arrange(dfshipping_copy,price_of_carrier)  

indices <- order(dfshipping_copy$price_of_carrier, na.last = FALSE, decreasing = FALSE)
tishipping_sort <- dfshipping_copy[indices,]
```

country + mean of price_of_carrier

```{r}
tishipping1 <- tishipping_sort %>% dplyr::group_by(country) %>% summarise(
mean_price_of_carrier = mean(price_of_carrier, na.rm=TRUE)) %>% dplyr::ungroup()
```

if we want to add this to our data that include mean of price_of_carrier

```{r}
  tishipping2 <- dfshipping_copy  %>% 
        group_by(country) %>% 
        summarise(Avg =mean(price_of_carrier,na.rm = TRUE))  %>%
        inner_join(dfshipping_copy)
```
