---
title: "Transformation"
author: "Bakro"
date: "10/27/2021"
output:
  html_document:
      toc: true
      toc_float: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dpylr package

***dplyr functions work with pipes and expect tidy data***

## look at !!

+---------------+-----------------------------------------------------+
| method        | description                                         |
+===============+=====================================================+
| rename()      | change columns names                                |
+---------------+-----------------------------------------------------+
| relocate()    | Move columns to new position.                       |
+---------------+-----------------------------------------------------+
| mutate()      | add cols,add cols \~operation ,delete cols          |
+---------------+-----------------------------------------------------+
| count()       | Count number of rows in each group                  |
|               |                                                     |
|               | defined by the variables in ... Also tally().       |
+---------------+-----------------------------------------------------+
| distinct()    | Remove rows with duplicate values.                  |
+---------------+-----------------------------------------------------+
| select()      | Extract columns as a table.                         |
+---------------+-----------------------------------------------------+
| slice()       | Select rows by position.                            |
+---------------+-----------------------------------------------------+
| filter()      | Extract rows that meet logical criteria.            |
+---------------+-----------------------------------------------------+
| arrange()     | Order rows by values of a column or columns (low to |
|               |                                                     |
|               | high), use with desc() to order from high to low.   |
+---------------+-----------------------------------------------------+
| rows_insert() | add/change/delete row or more rows                  |
|               |                                                     |
| rows_update() |                                                     |
|               |                                                     |
| rows_delete() |                                                     |
+---------------+-----------------------------------------------------+
| group_by()    | create a"grouped" copy of a table grouped by        |
|               | columns in ... dplyr                                |
|               |                                                     |
|               | functions will manipulate each "group" separately   |
|               | and combine                                         |
|               |                                                     |
|               | the results.                                        |
+---------------+-----------------------------------------------------+
| group_split() | filter data depend on group                         |
+---------------+-----------------------------------------------------+
| summarise()   | Apply summary functions to columns to create a new  |
|               | table ofsummary statistics. Summary functions take  |
|               | vectors as input fun: n,n\_                         |
|               | distinct,sum,mean,median,quantile,min,              |
|               |                                                     |
|               | max,sd,var                                          |
+---------------+-----------------------------------------------------+
| ungroup()     | filter data depend on conditions                    |
+---------------+-----------------------------------------------------+
| na_if()       | replace null valyes                                 |
+---------------+-----------------------------------------------------+
| sql()         | sql statements                                      |
+---------------+-----------------------------------------------------+
| pull()        | Extract column values as a vector, by name or       |
|               | index.                                              |
+---------------+-----------------------------------------------------+

## summarise()

+----------------------------------------------------------------+
| Apply summary functions to columns to create a new table of    |
|                                                                |
| summary statistics. Summary functions take vectors as input    |
| and                                                            |
|                                                                |
| return one value (sum,min,max,sd,.....)                        |
+----------------------------------------------------------------+

```{r warning=FALSE , message=FALSE}
library(tidyverse)
library(skimr)
```

```{r}
summarise(mtcars, avg = mean(mpg))
```

## count()

+-----------------------------------------------+
| Count number of rows in each group            |
|                                               |
| defined by the variables in ... Also tally(). |
+-----------------------------------------------+

```{r}
count(mtcars, cyl)
```

## group_by

+-------------------------------------------------------------------+
| create a"grouped" copy of a table grouped by columns in ... dplyr |
|                                                                   |
| functions will manipulate each "group" separately and combine     |
|                                                                   |
| the results.                                                      |
+-------------------------------------------------------------------+

```{r}
mtcars %>%
group_by(cyl) %>%
summarise(avg = mean(mpg))
```

## ungroup()

|                                 |
|---------------------------------|
| returns ungrouped copy of table |

```{r}
t <- mtcars %>%
group_by(cyl) %>%
mutate(mn = cyl/mean(cyl)) %>%
ungroup() 

head(t,10)
```

## Manipulate Cases

### filter()

|                                          |
|------------------------------------------|
| Extract rows that meet logical criteria. |

```{r}
filter(mtcars, mpg > 20)
```

### distinct()

|                                    |
|------------------------------------|
| Remove rows with duplicate values. |

```{r}
distinct(mtcars, gear)
```

### slice()

|                          |
|--------------------------|
| Select rows by position. |

```{r}
slice(mtcars, 10:15)
```

### slice_sample()

|                       |
|-----------------------|
| Randomly select rows. |

```{r}
slice_sample(mtcars, n = 5, replace = TRUE) # return 5 records
```

### slice_min() & slice_max()

|                                                 |
|-------------------------------------------------|
| Select rows with the lowest and highest values. |

```{r}
t <- slice_min(mtcars, mpg, prop = 0.25) # return 5 records
```

### arrange()

+-----------------------------------------------------+
| Order rows by values of a column or columns (low to |
|                                                     |
| high), use with desc() to order from high to low.   |
+-----------------------------------------------------+

```{r eval=FALSE}
arrange(mtcars, mpg)

arrange(mtcars, desc(mpg))
```

### add_row()

+-----------------------------------------------------------------------+
| Add one or more rows to a table.Select rows with the lowest and       |
| highest values.                                                       |
+-----------------------------------------------------------------------+

```{r eval=FALSE}
add_row(cars, speed = 1, dist = 1) 
```

## Manipulate Variables

Column functions return a set of columns as a new vector or table

### pull()

|                                                          |
|----------------------------------------------------------|
| **Extract column values as a vector, by name or index.** |

```{r}
t <- pull(mtcars, wt)
```

### select()

| Extract columns as a table. |
|-----------------------------|

```{r}
t <- select(mtcars, mpg, wt)
```

### relocate()

+-----------------------------------------------------+
| relocate(.data, ..., .before = NULL, .after = NULL) |
|                                                     |
| Move columns to new position.                       |
+-----------------------------------------------------+

```{r}
t<- relocate(mtcars, mpg, cyl, .after = last_col())
head(t,3)
```

+--------------------------------------------------------------------+
| ***Use these helpers with select() and across()***                 |
|                                                                    |
| ***e.g. select(mtcars, mpg:cyl)***                                 |
|                                                                    |
| ***contains(match) num_range(prefix, range) :, e.g. mpg:cyl***     |
|                                                                    |
| ***ends_with(match) all_of(x)/any_of(x, ..., vars) -, e.g,         |
| -gear***                                                           |
|                                                                    |
| ***starts_with(match) matches(match) everything()***               |
+--------------------------------------------------------------------+



## MANIPULATE MULTIPLE VARIABLES AT ONCE

> **across**(.cols, .funs, ..., .names = NULL) Summarise

or mutate multiple columns in the same way.

```{r}
summarise(mtcars, across(everything(), mean))
```

> **c_across**(.cols) Compute across columns in

row-wise data.

```{r}
#transmute(rowwise(UKgas), total = sum(c_across(1:2)))
```

## MAKE NEW VARIABLES

`Apply vectorized functions to columns. Vectorized functions take`

`vectors as input and return vectors of the same length as output.`

### mutate()

mutate(.data, ..., .keep = "all", .before = NULL,.after = NULL)

Compute new column(s). Also **add_column()**, **add_count()**, and
**add_tally()**.

```{r}
t<- mutate(mtcars, gpm = 1 / mpg)
head(t,3)
```

**transmute()**

transmute(.data, ...) Compute new column(s),drop others.

```{r}
t <-transmute(mtcars, gpm = 1 / mpg)
head(t,3)
```

### rename()

rename(.data, ...) Rename columns. Use rename_with() to rename with a
function.

```{r}
t <- rename(cars, distance = dist)
head(t,3)
```

## Row Names

Tidy data does not use rownames, which store a variable outside of the
columns. To work with the

rownames, first move them into a column.

**rownames_to_column()**

```{r}
t <- rownames_to_column(mtcars, var = "C")
head(t,3)
```

**column_to_rownames()**

```{r}
t <- column_to_rownames(t, var = "C")
head(t,3)
```

## Combine Tables

bind_cols(..., .name_repair) Returns tablesplaced side by side as a
single table. Column

lengths must be equal. Columns will NOT bematched by id (to do that look
at Relational Data

below), so be sure to check that both tables are ordered the way you
want before binding.

<hr>

bind_rows(..., .id = NULL)

Returns tables one on top of the other as a single table. Set .id to a
column name to

add a column of the original table names.

<hr>

Use a "Mutating Join" to join one table to columns from another,
matching values with the

rows that they correspond to. Each join retains a different combination
of values from the tables.

left_join(x, y, by = NULL, copy = FALSE,

suffix = c(".x", ".y"), ..., keep = FALSE,

na_matched = "na") Join matching

values from y to x.

right_join(x, y, by = NULL, copy = FALSE,

suffix = c(".x", ".y"), ..., keep = FALSE,

na_matches = "na") Join matching

values from x to y.

inner_join(x, y, by = NULL, copy = FALSE,

suffix = c(".x", ".y"), ..., keep = FALSE,

na_matches = "na") Join data. Retain

only rows with matches.

full_join(x, y, by = NULL, copy = FALSE,

suffix = c(".x", ".y"), ..., keep = FALSE,

na_matches = "na") Join data. Retain all

values, all rows.

## SET OPERATIONS

intersect(x, y, ...)

Rows that appear in both x and y.

setdiff(x, y, ...)

Rows that appear in x but not y.

union(x, y, ...)

Rows that appear in x or y.

(Duplicates removed). union_all()

retains duplicates.

Use setequal() to test whether two data sets

contain the exact same rows (in any order).

# RECAP

## 1- initialize data frames

```{r}
id <- c(1:10)

name <- c("John Mendes", "Rob Stewart", "Rachel Abrahamson", "Christy Hickman", "Johnson Harper", "Candace Miller", "Carlson Landy", "Pansy Jordan", "Darius Berry", "Claudia Garcia")

sex <- factor(c("M","F","M","F","M","F","M","F","M","M"))
age <- c(22,23,25,26,27,28,24,21,29,30)
children <- c(2,2,2,3,2,3,4,1,2,3)
people <- data.frame(id,name,sex,age,children)
```

```{r}
id <- c(1:5)
iscar <- c(TRUE,FALSE,TRUE,FALSE,TRUE)
distance_KM <- c(34.5,32,33,35,43)
car <- data.frame(id,iscar,distance_KM)
```

## 2- explore data

-   show in viewer

    ```{r}
    #View(people)
    #view(car)
    ```

-   structures of data frames

    ```{r}
     explore <- function(dataframe)
     {
     # is.data.frame() 
      class(dataframe) 
      mode(dataframe)
      length(dataframe)
      str(dataframe)
      skim(dataframe)
      summary(dataframe)
      nrow(dataframe)
      ncol(dataframe)
      attributes(dataframe)
      head(dataframe)
     }

    ```

    ```{r}
     explore(dataframe = people)
     explore(dataframe = car)
    ```

## 3- cleanData

-   rename columns

    ```{r}
    colnames(car) <- c("Id","Iscar","distance_km")
    colnames(car)
    ```

-   rename rows

    ```{r}
    row.names(car) <- c(str_c("zn",1:nrow(car)))
    row.names(car)
    ```

## 4- process data

-   **add columns**

    ```{r}

    people <- mutate(people,pow_of_age = age*age)
    head(people,3)
    ```

-   **delete columns**

    ```{r}
    people1 <- people 
    people1[,c(1:3)] <- NULL
    head(people1)
    ```

-   **sort**

    ```{r}
    head(arrange(people,desc(age)),3)
    head(arrange(people,age),3)
    ```

-   **indexing**

    ```{r}
    people1[1,1] # only one observation 
    slice(people1,1)
    people1[1:5,] # range of observation 
    samples <- sample(3, size=length(people), replace=TRUE, prob=c(.60,.20,.20)) # samplinoups g data to gr
    people1[samples == 1,c(1,2)] 
    slice_sample(people1,n=3)
    ```

-   **indexing variables inside dataframe**

    ```{r}
    people1$age[1:3]
    people1[["age"]][1:3]
    people1[c(1:3),1]
    ```

-   **selection and filtering**

    ```{r}
    subset(people1, select = age ,age >25); # select and filter
    ```

-   **spilting data**

    ```{r}
     splited_data <- split(people, f= people$sex, drop = TRUE)
    splited_data
    ```

## 5- dyplr

```{r}
dfBsp <- tibble(UID = 1:10, 
                Name = LETTERS[1:10], 
                Group = c("G1", "G3", "G2", 
                          "G3", "G1", "G2", 
                          "G2", "G3", "G1", "G2"), 
                Value = c(seq(100, 900, by = 100), NA))
dfBsp
```

```{r}
dfGBy <- dplyr::group_by(dfBsp, Group); dfGBy
```

```{r}
vlGrp <- is_grouped_df(dfGBy); vlGrp
```

```{r}
dfGBy <- dplyr::ungroup(dfGBy); dfGBy
vlGrp <- is_grouped_df(dfGBy); vlGrp
```

```{r}
liBspGrp <- dplyr::group_split(dplyr::group_by(dfBsp, Group)); liBspGrp
```

```{r}
names(liBspGrp) <- paste("Group ", unique(dfBsp$Group), sep = "")
liBspGrp
```

```{r}
dfBsp$Value <- dplyr::na_if(dfBsp$Value, 100); dfBsp
```

```{r}
dplyr::summarize(dplyr::group_by(dfBsp, Group), 
                 Anz = n(),
                 Min = min(Value, na.rm = TRUE),
                 Max = max(Value, na.rm = TRUE),
                 MW = mean(Value, na.rm = TRUE),
                 Med = median(Value, na.rm = TRUE),
                 StdAbw = sd(Value, na.rm = TRUE), 
                 Q25 = quantile(Value, na.rm = TRUE)["25%"], 
                 Q75 = quantile(Value, na.rm = TRUE)["75%"], 
                 AnzNAs = sum(is.na(Value)))
```

```{r}
dplyr::summarize(dplyr::group_by(dfBsp, Group), 
                 Anz = n(),
                 Min = min(Value, na.rm = TRUE),
                 Max = max(Value, na.rm = TRUE),
                 MW = mean(Value, na.rm = TRUE),
                 Med = median(Value, na.rm = TRUE),
                 StdAbw = sd(Value, na.rm = TRUE), 
                 Q25 = quantile(Value, na.rm = TRUE)["25%"], 
                 Q75 = quantile(Value, na.rm = TRUE)["75%"], 
                 AnzNAs = sum(is.na(Value)))
```
