---
title: "prepare case"
author: "Bakro"
date: "10/12/2021"
output: html_document
---

# prepare Case

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
print("This file was created within RStudio")
print("And now it lives on GitHub")
```

## deal with files

```{r eval=FALSE}
getwd()  # current working directory
ls()     # local object workspace
list.files() # files in current directory 
dir() # files in current directory 
?list.files   # help method
args(list.files) # argument in this method
old.dir <- getwd() # save current directory
dir.create("testdir") # create directory in current directory
setwd("testdir") # change directory
file.create("mytest.R") # create R file 
file.exists("mytest.R")  # if file exists in current directory 
file.info("mytest.R")  # info about fie size ,isdir,time ,exe
file.rename("mytest.R","mytest2.R") # rename file
file.copy("mytest2.R","mytest3.R") # copy 
file.path("mytest3.R")  # path of file
?dir.create             # help 
# create testdir2 and subfolder testdir3
dir.create(file.path("testdir2", "testdir3"), recursive = TRUE)
setwd(old.dir) # return to directory 

```

<div>

calculate required space for memory (remember basics)

</div>

**Reading data**

## load & save

```{r eval=FALSE}
load(url("http://dwoll.de/rKompakt/dat.Rdata"))    # datL, datW, subW

path <- paste0(getwd(), "/Woll2.Rdata")

save(datL, datW, subW, file = vcFile)

rm(datL, datW, subW)

path <- paste0(getwd(), "/Woll2.Rdata")

load(vcFile)
```

## download from internet

```{r eval=FALSE}
fileurl <- "..."
download.file(fileurl,destfile = "...", method = "curl") # https so that md :curl 
list.files("./data")
```

## source(), dump() bzw. dput(), dget()

```{r eval = FALSE}
dfData <- data.frame(PersID = paste("P", 1:4, sep = ""), 
                     Initials = c("AB", "CD", "EF", "GH"), 
                     Smoker = c("N", "N", "Y", "Y"), 
                     Age = c(10, 19, 11, 14))

# ... und ein weiteres Objekt:

string <- " welcome here"

vcFile = paste0(getwd(), "/Data_dump.R")
#con <- file(vcFile, "r")	

dump(c("string", "dfData"), file = vcFile)
#close(con)

rm(string, dfData)
## as module or library that include functions,data ,..
source(vcFile)

string
dfData
```

## dput() und dget()

```{r eval =FALSE}
dput(dfData)  # show structure of  Data.Frames

vcFile = paste(vcDir, "Data_dput.R", sep = "")

dput(dfData, file = vcFile)

dfDataNew <- dget(vcFile)

dfDataNew
```

## unstructure data

```{r eval =FALSE}
vcFile <- paste(vcDir, "TextData.txt", sep = "")
vcText <- readLines(vcFile)
vcText

# Inhalt eines Character-Vektors in einen Textfile schreiben
vcFile <- paste(vcDir, "TextDataCopy.txt", sep = "")
writeLines(vcText, vcFile, sep = "\n")
```

```{r eval =FALSE}
vcHTML <- readLines(url("https://www.infoq.com/news/", "r"))
vcFile <- paste(vcDir, "infoq.html")
writeLines(vcHTML,vcFile)
```

## read.table / read.csv

```{r eval =FALSE}
vcFile <- paste(vcDir, "TabData01.txt", sep = "")

dfDaten <- read.table(vcFile, 
                      header = TRUE, sep = ",", 
                      stringsAsFactors = TRUE )

vcFile <- paste(vcDir, "TabData02.txt", sep = "")

write.table(dfDaten, file = vcFile)
```

## Excel files

```{r eval = FALSE}
library(xlsx)
colIndex <- 2:3
rowIndex <- 1:4
data <- read.xlsx ("..." , sheetIndex = 1,header = TRUE
                   ,colIndex = colIndex ,rowIndex = rowIndex )
```

## Reading with DB (Mysql)

```{r eval =FALSE}
# sql reading   -----------------------------------------------------------------
# download packages and load it to programm
install.packages("DBI")
install.packages("RMySQL")
library(DBI)
library(RMySQL)
# first open mysql and change password usnig 
# ALTER USER 'root'@'localhost' IDENTIFIED WITH mysql_native_password BY 'root';


con <-  dbConnect(MySQL(), user = 'root', password = 'root', dbname = 'mydb',host = '127.0.0.1')
t <- dbListTables(con)  # return tables 
dbListFields(con,t[1]) # know fields name for table
query <- paste("select * from ",t[1])
rs <-  dbSendQuery(con, query )
data = fetch(rs, n=-1) # fetch data 
```

## Reading from HDF5

***see video !!***

## Reading from webpages

***see more for more info !!***

```{r eval =FALSE}
con = url("address")
htmlcode = readLines(con)
close(con)
```

## parsing with XML

more info in course 3 (R Getting and clean )

```{r eval =FALSE}
library(XML)
library(methods)
require(XML)
# https://www.w3schools.com/xml/simple.xml
result <- xmlParse(file = "data/simple.xml")
print(result)
 rootNode <- xmlRoot(result)
 names(rootNode)
 rootNode[[1]]
  rootNode[[1]][[1]]
xpathSApply(rootNode , "//name" ,xmlValue)
xpathSApply(rootNode , "//price" ,xmlValue)
```


## Reading json

```{r eval =FALSE}
library(jsonlite)
jsondata <- fromJSON("https://api.github.com/users/jtleek/repos")
names(jsondata)
names(jsondata$owner)
jsondata$owner$login


myjson <- toJSON(iris, pretty = TRUE)
cat(myjson)

iris5 <- fromJSON(myjson)
```

Reading From APIs

for example accessing Twitter from R and than convert json object

## reading from other packages

read.foo

```{r eval =FALSE}
read.arff(Weka)
read.dta(Stata)
read.mtp(Minitab)
read.octave(Octave)
read.spss(SPSS)
read.xport(SAS)
```

## read big data

```{r eval =FALSE}
library(data.table)
big_df <- data.frame(x =rnorm(1000000), y = rnorm(1000000))
file <- tempfile()
write.table(big_df,file = file ,row.names = FALSE ,sep = ":",quote = FALSE)
system.time(fread(file))
```

```{r eval =FALSE}
system.time(read.table(file,header = TRUE,sep = ":"))
```

```{r eval =FALSE}
t <- download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv")
```

## connection

```{r eval =FALSE}
vcFile <- paste(vcDir, "TextData.txt", sep = "")

conObj <- file(vcFile, "r")			# Open connection, read only
vcTextData <- readLines(conObj); vcTextData
close(conObj)					         # Close connectin

vcFile <- paste(vcDir, "TextDataCopy.txt")
conObj <- file(vcFile, "w")
writeLines(vcTextData, conObj)
close(conObj)
```

# Example

```{r eval =FALSE}
library(sqldf)
fileUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
download.file(fileUrl, destfile = "acs.csv")
acs <- read.csv("acs.csv")
head(acs)
#detach("package:RMySQL", unload=TRUE)
sqldf("select pwgtp1 from acs where AGEP < 50")
```

```{r}
htmlUrl <- url("http://biostat.jhsph.edu/~jleek/contact.html")
htmlCode <- readLines(htmlUrl)
close(htmlUrl)
c(nchar(htmlCode[10]), nchar(htmlCode[20]), nchar(htmlCode[30]), nchar(htmlCode[100]))
```
