---
title: "35_keras_Overfitting_and_Underfitting"
author: "Bakro"
date: "1/4/2022"
output: 
  html_document:
   toc: true
   toc_float: true
   toc_depth: 3
   theme: flatly
   highlight: zenburn
   df_print: paged
   code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overfitting and Underfitting

```{r warning=FALSE , message=FALSE}
library(keras)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tibble)
```

### Download the IMDB dataset

```{r}
num_words <- 1000
imdb <- dataset_imdb(num_words = num_words)

c(train_data, train_labels) %<-% imdb$train
c(test_data, test_labels) %<-% imdb$test
```

Rather than using an embedding as in the previous notebook, here we will multi-hot encode the sentences.
This model will quickly overfit to the training set.
It will be used to demonstrate when overfitting occurs, and how to fight it.

Multi-hot-encoding our lists means turning them into vectors of 0s and 1s.
Concretely, this would mean for instance turning the sequence [3, 5] into a 10,000-dimensional vector that would be all-zeros except for indices 3 and 5, which would be ones.

```{r}
multi_hot_sequences <- function(sequences, dimension) {
  multi_hot <- matrix(0, nrow = length(sequences), ncol = dimension)
  for (i in 1:length(sequences)) {
    multi_hot[i, sequences[[i]]] <- 1
  }
  multi_hot
}

train_data <- multi_hot_sequences(train_data, num_words)
test_data <- multi_hot_sequences(test_data, num_words)
```

```{r}
# Let’s look at one of the resulting multi-hot vectors. The word indices are sorted by frequency, so it is expected that there are more 1-values near index zero, as we can see in this plot:

first_text <- data.frame(word = 1:num_words, value = train_data[1, ])
ggplot(first_text, aes(x = word, y = value)) +
  geom_line() +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

```

### CREATE A BASELINE MODEL

```{r}
baseline_model <- 
  keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = num_words) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

baseline_model %>% compile(
  optimizer = "adam",
  loss = "binary_crossentropy",
  metrics = list("accuracy")
)

summary(baseline_model)
```

```{r}
baseline_history <- baseline_model %>% fit(
  train_data,
  train_labels,
  epochs = 20,
  batch_size = 512,
  validation_data = list(test_data, test_labels),
  verbose = 2
)
```

### CREATE A SMALLER MODEL

```{r}
smaller_model <- 
  keras_model_sequential() %>%
  layer_dense(units = 4, activation = "relu", input_shape = num_words) %>%
  layer_dense(units = 4, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

smaller_model %>% compile(
  optimizer = "adam",
  loss = "binary_crossentropy",
  metrics = list("accuracy")
)

summary(smaller_model)

smaller_history <- smaller_model %>% fit(
  train_data,
  train_labels,
  epochs = 20,
  batch_size = 512,
  validation_data = list(test_data, test_labels),
  verbose = 2
)
```

### CREATE A BIGGER MODEL

```{r}
bigger_model <- 
  keras_model_sequential() %>%
  layer_dense(units = 512, activation = "relu", input_shape = num_words) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

bigger_model %>% compile(
  optimizer = "adam",
  loss = "binary_crossentropy",
  metrics = list("accuracy")
)

summary(bigger_model)

bigger_history <- bigger_model %>% fit(
  train_data,
  train_labels,
  epochs = 20,
  batch_size = 512,
  validation_data = list(test_data, test_labels),
  verbose = 2
)
```

### PLOT THE TRAINING AND VALIDATION LOSS

```{r}
compare_cx <- data.frame(
  baseline_train = baseline_history$metrics$loss,
  baseline_val = baseline_history$metrics$val_loss,
  smaller_train = smaller_history$metrics$loss,
  smaller_val = smaller_history$metrics$val_loss,
  bigger_train = bigger_history$metrics$loss,
  bigger_val = bigger_history$metrics$val_loss
) %>%
  rownames_to_column() %>%
  mutate(rowname = as.integer(rowname)) %>%
  gather(key = "type", value = "value", -rowname)
  
ggplot(compare_cx, aes(x = rowname, y = value, color = type)) +
  geom_line() +
  xlab("epoch") +
  ylab("loss")
```

### Strategies

#### ADD WEIGHT REGULARIZATION

```{r}
l2_model <- 
  keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = num_words,
              kernel_regularizer = regularizer_l2(l = 0.001)) %>%
  layer_dense(units = 16, activation = "relu",
              kernel_regularizer = regularizer_l2(l = 0.001)) %>%
  layer_dense(units = 1, activation = "sigmoid")

l2_model %>% compile(
  optimizer = "adam",
  loss = "binary_crossentropy",
  metrics = list("accuracy")
)

l2_history <- l2_model %>% fit(
  train_data,
  train_labels,
  epochs = 20,
  batch_size = 512,
  validation_data = list(test_data, test_labels),
  verbose = 2
)
```

```{r}
compare_cx <- data.frame(
  baseline_train = baseline_history$metrics$loss,
  baseline_val = baseline_history$metrics$val_loss,
  l2_train = l2_history$metrics$loss,
  l2_val = l2_history$metrics$val_loss
) %>%
  rownames_to_column() %>%
  mutate(rowname = as.integer(rowname)) %>%
  gather(key = "type", value = "value", -rowname)
  
ggplot(compare_cx, aes(x = rowname, y = value, color = type)) +
  geom_line() +
  xlab("epoch") +
  ylab("loss")
```

As you can see, the L2 regularized model has become much more resistant to overfitting than the baseline model, even though both models have the same number of parameters

#### 

ADD DROPOUT

```{r}
dropout_model <- 
  keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = num_words) %>%
  layer_dropout(0.6) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dropout(0.6) %>%
  layer_dense(units = 1, activation = "sigmoid")

dropout_model %>% compile(
  optimizer = "adam",
  loss = "binary_crossentropy",
  metrics = list("accuracy")
)

dropout_history <- dropout_model %>% fit(
  train_data,
  train_labels,
  epochs = 20,
  batch_size = 512,
  validation_data = list(test_data, test_labels),
  verbose = 2
)
```

```{r}
compare_cx <- data.frame(
  baseline_train = baseline_history$metrics$loss,
  baseline_val = baseline_history$metrics$val_loss,
  dropout_train = dropout_history$metrics$loss,
  dropout_val = dropout_history$metrics$val_loss
) %>%
  rownames_to_column() %>%
  mutate(rowname = as.integer(rowname)) %>%
  gather(key = "type", value = "value", -rowname)
  
ggplot(compare_cx, aes(x = rowname, y = value, color = type)) +
  geom_line() +
  xlab("epoch") +
  ylab("loss")
```

dding dropout is a clear improvement over the baseline model.

To recap: here the most common ways to prevent overfitting in neural networks:

-   Get more training data.

-   Reduce the capacity of the network.

-   Add weight regularization.

-   Add dropout.

And two important approaches not covered in this guide are data augmentation and batch normalization.