---
title: "24_data_mining_association"
author: "Bakro"
date: "1/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Association Rule Mining

Association rules: apriori(), eclat() in package arules I Sequential patterns: arulesSequence I Visualisation of associations: arulesViz Function apriori()

Mine frequent itemsets, association rules or association hyperedges using the Apriori algorithm. The Apriori algorithm employs level-wise search for frequent itemsets. Default settings: I minimum support: supp=0.1 I minimum confidence: conf=0.8 I maximum length of rules: maxlen=10

```{r}
# The Titanic Dataset

load("./data/titanic.raw.rdata")
dim(titanic.raw)
## [1] 2201 4
idx <- sample(1:nrow(titanic.raw), 8)
titanic.raw[idx, ]

# find association rules with the APRIORI algorithm
library(arules)
rules <- apriori(titanic.raw, control=list(verbose=F),
parameter=list(minlen=2, supp=0.005, conf=0.8),
appearance=list(rhs=c("Survived=No", "Survived=Yes"),

default="lhs"))

# sort rules
quality(rules) <- round(quality(rules), digits=3)
rules.sorted <- sort(rules, by="lift")
# have a look at rules
inspect(rules.sorted)
#Rule #2 provides no extra knowledge in addition to rule #1,
#since rules #1 tells us that all 2nd-class children survived.
#I When a rule (such as #2) is a super rule of another rule (#1)
#and the former has the same or a lower lift, the former rule
#(#2) is considered to be redundant.
#I Other redundant rules in the above result are rules #4, #7
# and #8, compared respectively with #3, #6 and #5.


print(" Remove Redundant Rules ")
## find redundant rules
subset.matrix <- is.subset(rules.sorted, rules.sorted)
subset.matrix[lower.tri(subset.matrix, diag = T)] <- NA
redundant <- colSums(subset.matrix, na.rm = T) >= 1

## which rules are redundant
which(redundant)

## remove redundant rules
rules.pruned <- rules.sorted[!redundant]

## Remaining Rules
inspect(rules.pruned)

## interrupt result 
#Did children have a higher survival rate than adults?
#I Did children of the 2nd class have a higher survival rate than
#other children?
# Did children have a higher survival rate than adults?
# Did children of the 2nd class have a higher survival rate than
#other children?
# The rule states only that all children of class 2 survived, but
#provides no information at all about the survival rates of other
#classes.



# viZ
library(arulesViz)
# plot(rules.all)

plot(rules.sorted, method = "grouped")

plot(rules.sorted, method = "graph", control = list(type = "items"))

plot(rules.sorted, method = "paracoord", control = list(reorder = T))

```

### Association Rules

Association rules are rules presenting association or correlation between itemsets.\
support(A ⇒ B) = P(A ∪ B)\
confidence(A ⇒ B) = P(B\|A) =P(A ∪ B)/P(A)\
lift(A ⇒ B) = confidence(A ⇒ B)/P(B)=P(A ∪ B)/P(A)P(B)\

### Interestingness Measures

-   Which rules or patterns are the most interesting ones? One way is to rank the discovered rules or patterns withinterestingness measures.\

-   The measures of rule interestingness fall into two categories,subjective and objective[Freitas, 1998, Silberschatz and Tuzhilin, 1996].\

-   Objective measures, such as lift, odds ratio and conviction,are often data-driven and give the interestingness in terms of statistics or information theory.\

-   Subjective (user-driven) measures, e.g., unexpectedness and actionability, focus on finding interesting patterns by\
    matching against a given set of user beliefs.

```{r}
# https://www.kirenz.com/post/2020-05-14-r-association-rule-mining/
# create a list of baskets
market_basket <-  
  list(  
  c("apple", "beer", "rice", "meat"),
  c("apple", "beer", "rice"),
  c("apple", "beer"), 
  c("apple", "pear"),
  c("milk", "beer", "rice", "meat"), 
  c("milk", "beer", "rice"), 
  c("milk", "beer"),
  c("milk", "pear")
  )

# set transaction names (T1 to T8)
names(market_basket) <- paste("T", c(1:8), sep = "")

# packages
library(arules)
library(arulesViz)

trans <- as(market_basket, "transactions")
dim(trans)
itemLabels(trans)
summary(trans)
image(trans)
itemFrequencyPlot(trans, topN=10,  cex.names=1)

#Min Support 0.3, confidence as 0.5.
rules <- apriori(trans, 
                 parameter = list(supp=0.3, conf=0.5, 
                                  maxlen=10, 
                                  target= "rules"))

summary(rules)
inspect(rules)

# Viz
plot(rules)
plot(rules, measure = "confidence")
plot(rules, method = "two-key plot")
subrules <- head(rules, n = 10, by = "confidence")

plot(subrules, method = "graph",  engine = "htmlwidget")
```
