---
title: "25_data_mining_Social_Network_Analysis"
author: "Bakro"
date: "1/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Social Network Analysis  

```{r}
library(igraph)
# nodes
nodes <- data.frame(
name = c("Tom","Ben","Bob","John","Mary","Alice","Wendy","David"),
gender = c("M", "M", "M", "M", "F", "F", "F", "M"),
age = c( 16, 30, 42, 29, 26, 32, 18, 22)
)
# relations
edges <- data.frame(
from = c("Tom", "Ben", "Bob", "Mary", "Alice", "Wendy", "Wendy"),
to = c("John", "John", "John", "John","Mary", "Mary", "David")
)
# build a graph object
g <- graph.data.frame(edges, directed=F, vertices=nodes)

layout1 <- g %>% layout_nicely() ## save layout for reuse
g %>% plot(vertex.size = 30, layout = layout1)

## use blue for male and pink for female
colors <- ifelse(V(g)$gender=="M", "skyblue", "pink")
g %>% plot(vertex.size=30, vertex.color=colors, layout=layout1)
```

```{r}
# Graph Query
## nodes
V(g)
## edges
E(g)
## immediate neighbors (friends) of John
friends <- ego(g,order=1,nodes="John",mindist=1)[[1]] %>% print()
## female friends of John
friends[friends$gender == "F"]

```

```{r}
# Graph Query (cont.)
## 1- and 2-order neighbors (friends) of John
g2 <- make_ego_graph(g, order=2, nodes="John")[[1]]
g2 %>% plot(vertex.size=30, vertex.color=colors)
```

### Centrality Measures

I Degree: the number of adjacent edges; indegree and outdegree for directed graphs\
I Closeness: the inverse of the average length of the shortest paths to/from all other nodes\
I Betweenness: the number of shortest paths going through a node\

### Centrality Measures (cont.)

I Eigenvector centrality: the values of the first eigenvector of the graph adjacency matrix\
I Transivity, a.k.a clustering coefficient, measures the probability that the adjacent nodes of a node are connected.

```{r}
degree <- g %>% degree() %>% print()

closeness <- g %>% closeness() %>% round(2) %>% print()

betweenness <- g %>% betweenness() %>% print()

layout1 <- g %>% layout_nicely() ## save layout for reuse
g %>% plot(vertex.size = 30, layout = layout1)

eigenvector <- evcent(g)$vector %>% round(2) %>% print()

transitivity <- g %>% transitivity(type = "local") %>% print()

```

### Static Network Visualization

I Static network visualization I Fast in rendering big graphs I For very big graphs, the most efficient way is to save visualization result into a file, instead of directly to screen. I Save network diagram into files: pdf(), bmp(), jpeg(), png(), tiff()

```{r}
library(igraph)
## plot directly to screen when graph is small
plot(g)
## for big graphs, save visualization to a PDF file
pdf("mygraph.pdf")
plot(g)
graphics.off() ## or dev.off()
```

```{r}
library(visNetwork)
## Interactive Network Visualization
visIgraph(g, idToLabel=T) %>%
## highlight nodes connected to a selected node
visOptions(highlightNearest=T) %>%
## use different icons for different types (groups) of nodes
visGroups(groupname="person", shape="icon",
icon=list(code="f007")) %>%

## use FontAwesome icons
addFontAwesome() %>%
## add legend of nodes
visLegend() 
## to save to file
# %>% visSave(file = "network.html")
```

```{r}
library(visNetwork)
# V(g)$group <- V(g)$type
## visualization
data <- toVisNetworkData(g)
visNetwork(nodes=data$nodes, edges=data$edges) %>%
visGroups(groupname="tid",shape="icon",icon=list(code="f15c")) %>%
visGroups(groupname="person",shape="icon",icon=list(code="f007")) %>%
visGroups(groupname="addr",shape="icon",icon=list(code="f015")) %>%
visGroups(groupname="phone",shape="icon",icon=list(code="f095")) %>%
visGroups(groupname="email",shape="icon",icon=list(code="f0e0")) %>%
addFontAwesome() %>%
visLegend()
```
