---
title: "Shiny"
author: "Bakro"
date: "10/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Shiny

**Building an App**

-   A Shiny app is a web page (ui) connected to a computer running a live R session (server).
-   Users can manipulate the UI, which will cause the server to update the UI's displays (by running R code).

## Inputs

Collect values from the user.

Access the current value of an input object with

input\$\<inputId\>. Input values are reactive.

<div>

**actionButton**(inputId, label, icon,width, ...)

**actionLink**(inputId, label, icon, ...)

**checkboxGroupInput**(inputId, label, choices, selected, inline, width,

choiceNames, choiceValues)

**checkboxInput**(inputId, label, value,width)

**dateInput**(inputId, label, value, min,max, format, startview, weekstart,

language, width, autoclose,datesdisabled, daysofweekdisabled)

**dateRangeInput**(inputId, label, start,end, min, max, format, startview,weekstart

, language, separator, width,autoclose)

**fileInput**(inputId, label, multiple,accept, width, buttonLabel, placeholder)

**numericInput**(inputId, label, value,min, max, step, width)

**passwordInput**(inputId, label, value,width, placeholder)

**radioButtons**(inputId, label,choices, selected, inline, width,choiceNames, choiceValues)

**selectInput**(inputId, label, choices,selected, multiple, selectize, width, size)

Also selectizeInput()

**sliderInput**(inputId, label, min, max,value, step, round, format, locale, ticks,

animate, width, sep, pre, post,timeFormat, timezone, dragRange)

**submitButton**(text, icon, width)

(Prevent reactions for entire app)

**textInput**(inputId, label, value, width,

placeholder) Also textAreaInput()

</div>

```{r}
library(tidyverse)
library(shiny)
```

```{r letter-a, echo=FALSE}
learnr::question("What number is the letter A in the English alphabet?",
  answer("8"),
  answer("14"),
  answer("1", correct = TRUE),
  answer("23")
)
```

## Example 1

```{r}
# Define UI for app that draws a histogram ----
ui <- fluidPage(

  # App title ----
  titlePanel("Hello Shiny!"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Slider for the number of bins ----
      sliderInput(inputId = "bins",
                  label = "Number of bins:",
                  min = 1,
                  max = 50,
                  value = 30)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Histogram ----
      plotOutput(outputId = "distPlot")

    )
  )
)
# Define server logic required to draw a histogram ----
server <- function(input, output) {

  # Histogram of the Old Faithful Geyser Data ----
  # with requested number of bins
  # This expression that generates a histogram is wrapped in a call
  # to renderPlot to indicate that:
  #
  # 1. It is "reactive" and therefore should be automatically
  #    re-executed when inputs (input$bins) change
  # 2. Its output type is a plot
  output$distPlot <- renderPlot({

    x    <- faithful$waiting
    bins <- seq(min(x), max(x), length.out = input$bins + 1)

    hist(x, breaks = bins, col = "#75AADB", border = "white",
         xlab = "Waiting time to next eruption (in mins)",
         main = "Histogram of waiting times")

    })

}
shinyApp(ui, server)
```

## Example2

```{r}
library(shiny)

# Define UI for dataset viewer app ----
ui <- fluidPage(

  # App title ----
  titlePanel("Shiny Text"),

  # Sidebar layout with a input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Selector for choosing dataset ----
      selectInput(inputId = "dataset",
                  label = "Choose a dataset:",
                  choices = c("rock", "pressure", "cars")),

      # Input: Numeric entry for number of obs to view ----
      numericInput(inputId = "obs",
                   label = "Number of observations to view:",
                   value = 10)
    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Verbatim text for data summary ----
      verbatimTextOutput("summary"),

      # Output: HTML table with requested number of observations ----
      tableOutput("view")

    )
  )
)

# Define server logic to summarize and view selected dataset ----
server <- function(input, output) {

  # Return the requested dataset ----
  datasetInput <- reactive({
    switch(input$dataset,
           "rock" = rock,
           "pressure" = pressure,
           "cars" = cars)
  })

  # Generate a summary of the dataset ----
  output$summary <- renderPrint({
    dataset <- datasetInput()
    summary(dataset)
  })

  # Show the first "n" observations ----
  output$view <- renderTable({
    head(datasetInput(), n = input$obs)
  })

}

# Create Shiny app ----
shinyApp(ui = ui, server = server)
```

## Example3

```{r}
library(shiny)

# Define UI for slider demo app ----
ui <- fluidPage(

  # App title ----
  titlePanel("Sliders"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar to demonstrate various slider options ----
    sidebarPanel(

      # Input: Simple integer interval ----
      sliderInput("integer", "Integer:",
                  min = 0, max = 1000,
                  value = 500),

      # Input: Decimal interval with step value ----
      sliderInput("decimal", "Decimal:",
                  min = 0, max = 1,
                  value = 0.5, step = 0.1),

      # Input: Specification of range within an interval ----
      sliderInput("range", "Range:",
                  min = 1, max = 1000,
                  value = c(200,500)),

      # Input: Custom currency format for with basic animation ----
      sliderInput("format", "Custom Format:",
                  min = 0, max = 10000,
                  value = 0, step = 2500,
                  pre = "$", sep = ",",
                  animate = TRUE),

      # Input: Animation with custom interval (in ms) ----
      # to control speed, plus looping
      sliderInput("animation", "Looping Animation:",
                  min = 1, max = 2000,
                  value = 1, step = 10,
                  animate =
                    animationOptions(interval = 300, loop = TRUE))

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Table summarizing the values entered ----
      tableOutput("values")

    )
  )
)

# Define server logic for slider examples ----
server <- function(input, output) {

  # Reactive expression to create data frame of all input values ----
  sliderValues <- reactive({

    data.frame(
      Name = c("Integer",
               "Decimal",
               "Range",
               "Custom Format",
               "Animation"),
      Value = as.character(c(input$integer,
                             input$decimal,
                             paste(input$range, collapse = " "),
                             input$format,
                             input$animation)),
      stringsAsFactors = FALSE)

  })

  # Show the values in an HTML table ----
  output$values <- renderTable({
    sliderValues()
  })

}

# Create Shiny app ----
shinyApp(ui, server)
```

## Example4

```{r}
library(shiny)

# Define UI for random distribution app ----
ui <- fluidPage(

  # App title ----
  titlePanel("Tabsets"),

  # Sidebar layout with input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Select the random distribution type ----
      radioButtons("dist", "Distribution type:",
                   c("Normal" = "norm",
                     "Uniform" = "unif",
                     "Log-normal" = "lnorm",
                     "Exponential" = "exp")),

      # br() element to introduce extra vertical spacing ----
      br(),

      # Input: Slider for the number of observations to generate ----
      sliderInput("n",
                  "Number of observations:",
                  value = 500,
                  min = 1,
                  max = 1000)

    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Tabset w/ plot, summary, and table ----
      tabsetPanel(type = "tabs",
                  tabPanel("Plot", plotOutput("plot")),
                  tabPanel("Summary", verbatimTextOutput("summary")),
                  tabPanel("Table", tableOutput("table"))
      )

    )
  )
)

# Define server logic for random distribution app ----
server <- function(input, output) {

  # Reactive expression to generate the requested distribution ----
  # This is called whenever the inputs change. The output functions
  # defined below then use the value computed from this expression
  d <- reactive({
    dist <- switch(input$dist,
                   norm = rnorm,
                   unif = runif,
                   lnorm = rlnorm,
                   exp = rexp,
                   rnorm)

    dist(input$n)
  })

  # Generate a plot of the data ----
  # Also uses the inputs to build the plot label. Note that the
  # dependencies on the inputs and the data reactive expression are
  # both tracked, and all expressions are called in the sequence
  # implied by the dependency graph.
  output$plot <- renderPlot({
    dist <- input$dist
    n <- input$n

    hist(d(),
         main = paste("r", dist, "(", n, ")", sep = ""),
         col = "#75AADB", border = "white")
  })

  # Generate a summary of the data ----
  output$summary <- renderPrint({
    summary(d())
  })

  # Generate an HTML table view of the data ----
  output$table <- renderTable({
    d()
  })

}

# Create Shiny app ----
shinyApp(ui, server)
```
