---
title: "Machine_Learning_project"
author: "Bakro"
date: "11/12/2021"
output: 
  html_document:
   toc: true
   toc_float: true
   toc_depth: 3
   theme: flatly
   highlight: zenburn
   df_print: paged
   code_folding: hide 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:

[[https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv]{.ul}](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv)

The test data are available here:

<https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>

```{r}
library(data.table)
library(skimr)
library(lattice)
library(ggplot2)
library(caret)
library(kernlab)
library(rattle)


# download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv" , destfile = "./data/pml_training.csv" , method = "curl")
pml_training <- fread("./data/pml-training.csv")

# download.file(url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv" , destfile = "./data/pml_testing" ,method = "curl")
pml_testing <- fread("./data/pml-testing.csv")

```

```{r}

# structure of data 
# skim_without_charts(pml_training)
# missing values in our data 
# propNAs <- colMeans(is.na(pml_training))
# table(propNAs)
# unnecessary data selecetion interested data for this project  
# remove columns
pml_training <- pml_training[,-c(1:7)]
pml_testing <- pml_testing[,-c(1:7)]
EmptyCols <- which(colSums(is.na(pml_training) |pml_training=="")>0.9*dim(pml_training)[1]) 

pml_training <- pml_training[,-..EmptyCols]
pml_testing <- pml_testing[,-..EmptyCols]
pml_training$classe <- as.factor(pml_training$classe)

# do same thing with testing data
pml_testing$classe <- as.factor(pml_testing$classe)
```

## Data slicing

```{r}
inTrain <- createDataPartition(y = pml_training$classe, p = 0.7, list = FALSE)

training <- pml_training[inTrain, ]

crossval <- pml_training[-inTrain, ]


```

## Preprocessing Data

```{r}
# Removing Zero Covariates
t <- nearZeroVar(training,saveMetrics=TRUE)
# when nzv = TRUE, those variables should be thrown out
# take indices of nzv = false 
indices <- which(!t$nzv)
training <- training[, ..indices ]
crossval <- crossval[, ..indices ]
# do same thing with testing data
pml_testing <- pml_testing[, ..indices ]

index_of_outcome <- which(colnames(training) == "classe")
# from skim we found there is missing values >> impute and standard data
# create preProcess object for all predictors (index_of_output = outcome)
preObj <- preProcess(training[,-..index_of_outcome],method=c("knnImpute","center","scale"))

# apply preprocessing on  training set
traindata <- predict(preObj,training[,-..index_of_outcome])
traindata$classe <- training$classe
# apply preprocessing on  cross validation 
cvdata <- predict(preObj,crossval[,-..index_of_outcome])
cvdata$classe <- crossval$classe
# normalize test set using training parameters
testdata <- predict(preObj,pml_testing)
rm(pml_testing ,pml_training)
```

## modals
### Decision Tree
```{r}
# set train control 
control <- trainControl(method="cv", number=3, verboseIter=F)
# first important thing our model is classification non linear model 
table(traindata$classe)
# Prediction with  Decision Tree
Fit_decision_Tree <-train(classe ~ .,method="rpart",data=traindata ,trControl = control,                     tuneLength = 5)
rattle::fancyRpartPlot(Fit_decision_Tree$finalModel)
testPC <-  predict(Fit_decision_Tree ,cvdata[,-..index_of_outcome])

# compare results
mrpart <-confusionMatrix(cvdata$classe,testPC) # ACC : 0.6615 

```
### random_Forest
```{r}
# take sample and test on than take total data

Fit_random_Forest <- train(classe ~ ., data = traindata, method = "rf" ,trControl = control, tuneLength = 5)
testPC <-  predict(Fit_random_Forest ,cvdata[,-..index_of_outcome])
mrf <- confusionMatrix(cvdata$classe,testPC) # ACC : 0.9997 

# A generic method for calculating variable importance for objects produced by train and method specific meth
# ?varImp
vi <- varImp(Fit_random_Forest)$importance



```
### boosting for part of data
```{r}
## here i have apply boosting only for section of the data
set.seed(100)
sampling <- sample(nrow(traindata),100)



# boosting 
Fit_boosting <- train(classe ~ ., method="gbm",data=traindata[sampling,],verbose=FALSE
                ,trControl = control, tuneLength = 5)

testPC <-  predict(Fit_boosting ,cvdata[,-..index_of_outcome])
mboosting <-confusionMatrix(cvdata$classe,testPC) # ACC : 0.9988
```


**S**upport **V**ector **M**achine

```{r}

mod_svm <- train(classe~., data=traindata, method="svmLinear", trControl = control, tuneLength = 5, verbose = F)
pred_svm <- predict(mod_svm, cvdata)
msvm <- confusionMatrix(pred_svm, factor(cvdata$classe))




```

## RESULTS

```{r}
# combine accuracy of this models 
# compare between them 
# i have take sampling for some models 
models <- c("Tree", "Rf", "Gbm/boosting","SVM")
accuracy <- round(c( mrpart$overall[1], mrf$overall[1], mboosting$overall[1] ,msvm$overall[1]),3) #accuracy
oos_error <- 1 - accuracy #out of sample error
data.frame(accuracy = accuracy, oos_error = oos_error, row.names = models)
```

## predict on testingset
```{r}
## Tree
pred_test_rpart <-  predict(Fit_decision_Tree ,testdata)
## RF
pred_test_rf <-  predict(Fit_random_Forest ,testdata)
## Gbm/boosting
testPC <-  predict(Fit_boosting ,testdata)
## svm
pred_test_svm <- predict(mod_svm, testdata)
```

### PCA

```{r}
# mod_pca <- train(classe~., data=traindata, method="rpart",preProcess="pca", trControl = control, tuneLength = 5)
# pred_pca<- predict(mod_pca, cvdata)
# mpca <- confusionMatrix(pred_pca, factor(cvdata$classe))

```

