---
title: "Air_pollution"
author: "Bakro"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{html}
<style>

div.red { background-color:#ff00ff; border-radius: 5px; padding: 20px;}
div.green { background-color:#00ff00; border-radius: 5px; padding: 20px;}
span.tag {color:crimson; border-radius: 10px; font-weight: bold;
background: #5bc0de88 ;padding: 5px; }
</style>
```
</style>

# Air pollution

**tags:** [importing data]{.tag} [protected /test case]{.tag} [handlingerror]{.tag} [optimize reading file]{.tag} [deal with multifiles]{.tag} [documentation]{.tag}

[dataframe]{.tag} [optimize memory]{.tag} [dp algorithm]{.tag} [control statement]{.tag} [debug]{.tag} [test]{.tag} [missingvalues]{.tag} [profoling]{.tag} [loops]{.tag}

::: {.alert .alert-block .alert-info style="font-size:14px; font-family:verdana; line-height: 1.7em;"}
📌   in this tutorial we will go from worse situation where we can detect errors and debug with bad performance for memory and time to good situation (especially when you use this fun more and more) but when you use this app for only time it's not a big deal to find optimizing solution

i have made some bugs for this tutorial
:::

## skim your app:

write algorithms that achieve your work than divide your problem to many

small problems that connect together (divide and conquer ) easier to test and read ...

(add function and add comments to this function and than try to convert this to code)

```{r eval=FALSE}
# this function find mean of  pollution depend on his type during 
# period of time (ids of files)

# params 
# directory : where do you find data ?
# pollutant : pollution art (sulfate/nitrate)
# id : contain id of files that will be monitor 

pollutantmean0 <- function(directory, pollutant ,  id = 1:332){
  # check data 

  # combine files
  
  # find mean  drop NA
  
  
} # end function polluntantmean 
```

## Version1

```{r eval=FALSE}
pollutantmean1 <- function(directory, pollutant ,  id = 1:332){
  # check data 
  if(missing(directory) ) stop("you should write directory")
  if(missing(pollutant) ) stop("you should write pollutant")
  if (pollutant != sulfate | pollutant != nitrate){ 
    stop("pollutant should be sulfate or nitrate")
  }
  # check numbers if in range 1 to 332
  
} # end function polluntantmean 

#  group1 
pollutantmean0("specdata","sulfate",1)
pollutantmean0("df","sulfate",1)
#  group2 
pollutantmean0("specdata","sulfate",1)
pollutantmean0("specdata","nitrate",1)
pollutantmean0("specdata","cbb",1)
# group3 
pollutantmean0("specdata","sulfate",-1)  # negative value
pollutantmean0("specdata","sulfate",-1:-3)  # range negative value
pollutantmean0("specdata","sulfate",c(1,0))
pollutantmean0("specdata","sulfate",c(1,-5))
pollutantmean0("specdata","sulfate",c(1:10))
pollutantmean0("specdata","sulfate",c(1,5))
pollutantmean0("specdata","sulfate",c(-5,5))
pollutantmean0("specdata","sulfate",c(320:335))


## make test situation in this stage verify your code 
## easier to find bug when you have error it is easier to fix 
#if (pollutant != "sulfate" & pollutant != "nitrate")
```

## Version2

```{r eval=FALSE}
pollutantmean0 <- function(directory, pollutant ,  id = 1:332){
  # check data 

  # define df
  # over a loop 
  # read file 
  # what happened if this file not existed 
  # detect classes of this table so that you can read quicker next time (only one)
  # combine data to df 
   data <- data.frame()
  firsttime <- TRUE
  # read file  1>> 001  , 91 >> 091 
    # directory <- "specdata"
    # numfile <- formatC(ind,width = 3,flag = "0")
    # vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
    # read.csv(vcFile)
  for (ind in id) {
    numfile <- formatC(ind,width = 3,flag = "0")
    vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
    # check if file exists
    if(!file.exists(vcFile)){
      warning(paste(vcFile , "this file doesn't esist"))
      next
    }
    # open connection 
    con <- file(vcFile,"r") 
    if(firsttime){
      temp <- read.csv(con)
      classes <- sapply(temp,class)
    }else{
      temp <- read.csv(con,colClasses = classes)
    }
    
    close(con)
    # combine data
    if(colnames(data) == colnames(temp)){
      data <- rbind(data,temp)} else{
        warning(paste(vcFile,"have another columns"))
      }
    
  } # end for 
  print("------------------------")
  
  mean <-  data %>%
    summarise(mean(.data[[pollutant]],na.rm =TRUE))
  return(paste(mean))
  print("------------------------")
  
  
} # end function polluntantmean


pollutantmean0("specdata","sulfate",1)
debug(pollutantmean0)
# 3-  apply  instead of tapply 
# 4-  sapply/lapply instead of tapply
# 5-  Error in if (colnames(data) == colnames(temp)) { : 
      #argument is of length zero
# 6-  is.null(data) doesn't work
```

## version3

```{r}
pollutantmean_v3 <- function(directory, pollutant ,id = 1:332){
  # check data 
  if(missing(directory) ) stop("you should write directory")
  if(missing(pollutant) ) stop("you should write pollutant")
  if (pollutant != "sulfate" & pollutant != "nitrate"){
    stop("pollutant should be sulfate or nitrate")
  }
  # check range of files 
  data <- NULL  # data <- data.frame()
  firsttime <- TRUE
for (ind in id) {
    numfile <- formatC(ind,width = 3,flag = "0") # sprintf("%03d", monitor)
    vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
    # check if file exists
    if(!file.exists(vcFile)){
      warning(paste(vcFile , "this file doesn't esist"))
      next
    }
    # open connection 
    con <- file(vcFile,"r") 
    if(firsttime){
      temp <- read.csv(con)
      classes <- lapply(temp,class)
      data <- temp
      firsttime <- FALSE
      close(con)
      next
    }else{
      temp <- read.csv(con,colClasses = classes)
     #temp <- read.csv(con) #  comment 2
      #temp<- fread(vcFile)
    }
    
    close(con)
    # combine data
    
    if(identical(colnames(data) , colnames(temp)) ){ # !is.null(data) &
    data <- rbind(data,temp)} else{
      warning(paste(vcFile,"have another columns"))
    }
      
} # end for 
print("------------------------")

  mean <-  data %>%
    summarise(mean(.data[[pollutant]],na.rm =TRUE))
  return(paste(mean))
  print("------------------------")

  
} # end function polluntantmean
pollutantmean("specdata","sulfate",1:10)
# group3 
pollutantmean("specdata","sulfate",-1)  # negative value
pollutantmean("specdata","sulfate",-1:-3)  # range negative value
pollutantmean("specdata","sulfate",c(1,0))
pollutantmean("specdata","sulfate",c(1,-5))
pollutantmean("specdata","sulfate",c(1:10))
pollutantmean("specdata","sulfate",c(1,5))
pollutantmean("specdata","sulfate",c(-5,5))
pollutantmean("specdata","sulfate",c(320:335))
#what happened if we haven't deal with missing values ?
debug(pollutantmean)
undebug(pollutantmean)
```

## Version4

```{r}
pollutantmean_v4<- function(directory, pollutant ,id = 1:332){
  # check data 
  if(missing(directory) ) stop("you should write directory")
  if(missing(pollutant) ) stop("you should write pollutant")
  if (pollutant != "sulfate" & pollutant != "nitrate"){
    stop("pollutant should be sulfate or nitrate")
  }
  
  
  tryCatch(
    {
      # check range of files 
      data <- NULL  # data <- data.frame()
      firsttime <- TRUE
      for (ind in id) {
        numfile <- formatC(ind,width = 3,flag = "0") # sprintf("%03d", monitor)
        vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
        # check if file exists
      
        # open connection 
        con <- file(vcFile,"r") 
        if(firsttime){
          temp <- read.csv(con)
          classes <- lapply(temp,class)
          data <- temp
          firsttime <- FALSE
          close(con)
          next
        }else{
          #temp <- read.csv(con,colClasses = classes)
          temp <- read.csv(con)
        }
        
        close(con)
        # combine data
        
        if(identical(colnames(data) , colnames(temp)) ){ # !is.null(data) &
          data <- rbind(data,temp)} else{
            warning(paste(vcFile,"have another columns"))
          }
        
      } # end for 
    }, # end trycatch exp
    error = function(e) {errorCondition("errror: ", e)},
    warning = function(w) {warning("Warning : ", w)}, 
    finally = { cat("finally  \n") }
  )
 
  print("------------------------")
  
  mean <-  data %>%
    summarise(mean(.data[[pollutant]],na.rm =TRUE))
  return(paste(mean))
  print("------------------------")
  
  
} # end function polluntantmean
pollutantmean_v4("specdata","sulfate",-1)  # negative value
pollutantmean_v4("specdata","sulfate",-1:-3)  # range negative value
pollutantmean_v4("specdata","sulfate",c(1,0))
pollutantmean_v4("specdata","sulfate",c(1,-5))
pollutantmean_v4("specdata","sulfate",c(1:10))
pollutantmean_v4("specdata","sulfate",c(1,5))
pollutantmean_v4("specdata","sulfate",c(-5,5))
pollutantmean_v4("specdata","sulfate",c(320:335))

# what difference between trycatch and control statements 
# when we use try catch 
```

## Version5

```{r}
pollutantmean_v5 <- function(directory, pollutant ,id = 1:332){
  # check data
  
  
  # define vector that include values(!missing ) 
  means <- c()
  
  for(monitor in id){
    # get path with standard format 00x,0xx,xxx
    path <- paste(getwd(), "/data/", directory, "/", sprintf("%03d", monitor), ".csv", sep = "")
    # read file 
    monitor_data <- read.csv(path)
   
    # select interested data
    interested_data <- monitor_data[pollutant]
    # remove missing value and add it to vector
    means <- c(means, interested_data[!is.na(interested_data)])
  }
  # calculate mean 
  mean(means)
  
} # end function polluntantmean 
```

## Version6

```{r}
# first time (10,20) (20,30)
result <- data.frame(filenumber= integer(),Avg =numeric())
pollutantmean_v6 <- function(directory , pollutant ,id = 1:332){

  # ind <- 1
  # directory = "specdata"
  # pollutant ="sulfate"
   means <- c()
  for (ind in id) {
    path <- paste(getwd(), "/data/", directory, "/",  sprintf("%03d", ind)
                  , ".csv", sep = "")
    data <- read.csv(path)
    interested_data <-  data[pollutant] 
    mean <-  mean(interested_data[!is.na(interested_data)])
    result <<- rbind(result,data.frame(filenumber =ind,Avg = mean))
  } # end for

} # end function 
pollutantmean_v6("specdata","sulfate",c(1:10))
result
undebug(pollutantmean04)
   pollutantmean04("specdata","sulfate",c(1,5))  
```

## profling

```{r}
Rprof(line.profiling=TRUE)
pollutantmean04("specdata", "sulfate")
 Rprof(NULL)
 summaryRprof()
```

## performance

[time]{.tag} [memory]{.tag} [accuracy]{.tag} [ui]{.tag} [services]{.tag}

### V3

```{r}
# v3
# reading with & without colnames
 start <- proc.time()
 pollutantmean_v3("specdata", "sulfate")
 end <- proc.time()
 print(end - start)
# user  system elapsed 
# 11.09    3.03   14.31 
# user  system elapsed  without 
# 10.35    3.12   13.68
library(data.table) 
 # in this example it's not a big deal (9.58 ,2.95,12.90)
 # than where most time go ? merge data 
 
```

### V5

```{r}
 start <- proc.time()
 pollutantmean_v5("specdata", "sulfate")
 end <- proc.time()
 print(end - start)
# user  system elapsed 
# 1.03    0.09   1.21 

```

V6

```{r}
 start <- proc.time()
 pollutantmean_v6("specdata", "sulfate")
 end <- proc.time()
 print(end - start)
# user  system elapsed 
# 1.03    0.09   1.28
 # but when you will have 
 mean(result[1:332,2])
```

## complete

```{r}
# pollutantmean(directory, pollutant,id = 1:332)
complete <- function(directory, id = 1:10){
  results <- data.frame(id=numeric(0), nobs=numeric(0))
  for (ind in id) {
    numfile <- formatC(ind,width = 3,flag = "0")
    vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
    # open connection 
    con <- file(vcFile,"r")
    temp <- read.csv(con)
    close(con)
    
    co <-  temp %>% 
      filter(!is.na(sulfate) & !is.na(nitrate)) %>%
      count()
    #print(paste( numfile ," ", co , " \n"))
    results <- rbind(results, data.frame(id=ind, nobs=co))
  } # end for 
  return(results) 
  
  
} # end function polluntantmean 
```

## correlation

```{r}
# cor between two factors
corr <- function(directory,threshold =0){
  #data <- data.frame()
  correlation <- NULL
  
  for (ind in 1:332) {
    numfile <- formatC(ind,width = 3,flag = "0")
    vcFile <- paste(getwd() ,"/data/",directory,"/", numfile , ".csv", sep = "")
    # open connection 
    con <- file(vcFile,"r")
    temp <- read.csv(con)
    close(con)
    
    temp<-  temp %>% 
      filter(!is.na(sulfate) & !is.na(nitrate)) 
    if(nrow(temp) > threshold)
      correlation <- c(correlation,cor(temp$sulfate,temp$nitrate))
    
  } # end for 
  return(correlation)
} 
cr <- corr(vcDir)
summary(cr)



```
