---
title: "graphics_1"
author: "Bakro"
date: "1/1/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)

```
## Basics

### curves

curve(function,from,to)

draw curve for sin,cos,exp,log,...

```{r include=FALSE}
#?curve     # help
curve(x^2-4, from = -2, to = +2)

```

type of curves

+---------+----------------------------+-----------+--------------------+
| abk     | type                       | abk       | type               |
+=========+============================+===========+====================+
| p       | points                     | s         | stairs curve       |
+---------+----------------------------+-----------+--------------------+
| l       | line                       | h         | verticals lines    |
+---------+----------------------------+-----------+--------------------+
| b       | separate line and points   | n         | no data            |
+---------+----------------------------+-----------+--------------------+
| o       | overlapped line and points |           |                    |
+---------+----------------------------+-----------+--------------------+

```{r}
curve(x^2-1, from = -2, to = +2, type = "l")  # Durchgehende Linien
 
```

### plots

```{r}
graphics.off()
vnX <- seq(from = -2, to = 2, by = 0.1);  
vnY <- vnX^2 - 1; 
plot(vnX, vnY, type = "p") # type same previous
```

pch shape

| num | shape            | num | shape               |
|-----|------------------|-----|---------------------|
| 1   | unfilled circles | 8   | stars               |
| 2   | triangles        | 16  | filled circles      |
| 3   | \+               | 20  | small filled circle |
| 4   | x                |     |                     |

```{r}
# make tables 

plot(vnX, vnY, type = "p", pch = 0 , xlab = "X Label", 
     ylab = "Y Label",
     xlim = c(-2, +2), 
     ylim = c(0, 10),
     main = "Blank Plotting Canvas")      
     
```

```{r eval = FALSE}
mnX <- c(1:9)
dim(mnX) <- c(3,3)
mnY <- c(10:18)
dim(mnY) <- c(3,3)
matplot(mnX, mnY, type = "l")  # matrix 
```

```{r}
plot( as.factor(diamonds$color)) # freq of values in factor bargraph 
```

### save graph

see plot tab

```{r eval=FALSE}
graphics.off()
dir <- getwd()
pdf(file = paste(dir,"/Grafik02.pdf" ,sep = ""))
curve(x^2-1, from = -2, to = +2)

dev.off()  
```

### multiplots

```{r}
par(mfrow = c(2, 2))				# 2 row,col
curve(1*x, from = -2, to = +2); curve(x^2, from = -2, to = +2)
curve(x^3, from = -2, to = +2); curve(x^4, from = -2, to = +2)
```


```{r}
#file.remove("Grafik03.pdf")
pdf(file = paste(getwd(),"/Grafik03.pdf" ,sep = ""))	
par(mfrow = c(2, 2))				# 2 row,col
curve(1*x, from = -2, to = +2); curve(x^2, from = -2, to = +2)
curve(x^3, from = -2, to = +2); curve(x^4, from = -2, to = +2)

```

### colors

```{r}

#colors()
palette()
rgb(0.5, 1.0, 0.75)
rgb(50, 100, 75, maxColorValue = 255)

```

```{r}
 plot.new()
 plot.window(xlim=c(0,1), ylim=c(5,10))
 #abline(a=3, b=1)
 curve(x^2,0,1)
 axis(1)
 axis(2)
 legend(x = "topright", legend = c("x/^2", "x"),
       col = c("blue", "red"), lty = c(1, 2))
 title(main="The Overall Title")
 title(xlab="An x-axis label")
 title(ylab="A y-axis label")
 grid(col = "grey", lty = 4, lwd = 1)
segments(x0 = -1, y0 = +1, x1 = 1, y1 = 3,
         col = "blue", lty = 1, lwd = 2)
 box()
```


```{r}
plot.new()
abline(a = 2, b = 0.5, col = "green", lty = 1, lwd = 2)
abline(h = 3, col = "darkgreen", lty = 2, lwd = 2)
abline(v = 0, col = "orange", lty = 2, lwd = 2)
lines(vnX, vnY, col = "brown", lty = 3, lwd = 2)
```

```{r}
vnX <- seq(from = -2, to = 2, by = 0.1); 
vnY <- vnX^2 - 1; 
plot(vnX, vnY, type = "n")  # Koordinatensystem etablieren

points(vnX, vnY)
lines(vnX, vnY)
```

```{r}
plot(vnX, vnY, type = "l", col = "red", lwd = 2, frame = FALSE)
box(which = "figure")
box(which = "plot")
```

```{r}
graphics.off() 
plot(vnX, vnY, type = "n")      # keine Daten im Diagramm
rect(c(-2, -1, 0, 1), c(-1, -1, -1, -1), 
     c(-1, 0, 1, 2), c(3, 3, 3, 3), 
     border = NULL, col = c("yellow", "orange", "red", "purple"))

lines(vnX, vnY, col = "blue", lty = 1, lwd = 2)
```

```{r}
plot(vnX, vnY, type = "l", col = "red", lwd = 2)
polygon(x = c(-1, 0, 1), y = c(1, 2, 1))
polygon(x = c(-2, -1, 0, 0, -2), y = c(0, 2, 1, 0, -1))
```



## ggEdit (uI)

```{r eval=FALSE}
#install.packages("ggedit")
library(ggedit)

plLenWid <- ggplot2::ggplot(iris,ggplot2::aes(x =Sepal.Length,y=Sepal.Width))

plLenWid <- plLenWid + 
   ggplot2::geom_point(ggplot2::aes(colour=Species)) + 
   ggplot2::geom_line()

if(interactive()){      # interactive()=TRUE, falls R interaktiv genutzt wird
   ## Not run: 
   plLenWid_new <- ggedit:ggedit(plLenWid)
   plLenWid_new
   
   ## End(Not run)
}

```

```{r eval=FALSE}
#install.packages("ggplotgui")
library(ggplotgui)

# Aufruf ohne Data.Frame / Tibble
ggplotgui::ggplot_shiny()

# Aufruf mit Data.Frame / Tibble
ggplotgui::ggplot_shiny(diamonds)
```

```{r eval=FALSE}
#install.packages("esquisse")
library(esquisse)
library(ggplot2)
# Aufruf ohne Data.Frame / Tibble
esquisse::esquisser()

ggplot(diamonds) +
 aes(x = carat, y = price, colour = cut) +
 geom_point(shape = "circle", size = 3.45) +
 scale_color_hue(direction = 1) +
 labs(title = "carat Vs price") +
 theme_gray() +
 theme(legend.position = "none")

esquisse::esquisser(diamonds)

# ... oder auch im Browser
esquisse::esquisser(diamonds, viewer = "browser")
```