---
title: "graphics_2"
author: "Bakro"
date: "1/1/2022"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggExtra)
```

## ggplot

### Bascis_shapes

```{r}
a <- ggplot(data =economics[1:10,] , aes(date,unemploy))
b <- ggplot(data =seals[1:10,] , aes(x = long,y = lat))
```

### blank & curve

```{r}
## geom blank 
g1 <- a+geom_blank() +ggtitle("blank")
g2 <- b + geom_curve(aes(yend = lat + 1,
xend = long + 1), curvature = 1) +ggtitle("curve")
```

### path &polygon

```{r}
## geom path 

g3 <-a + geom_path(lineend = "butt",
linejoin = "round", linemitre = 1) +ggtitle("path")

## geom polygon 
g4 <- a + geom_polygon(aes(alpha = 50))+ggtitle("polygon")
```

### rectangle & rippon

```{r}
## geom rect 
g5 <- b + geom_rect(aes(xmin = long, ymin = lat,
xmax = long + 1, ymax = lat + 1)) + ggtitle("rectangle")
```

```{r}
gridExtra::grid.arrange(g1,g2,g3,g4,g5,nrow = 2)
```
### LINE SEGMENTS

```{r}
g1 <- b + geom_abline(aes(intercept = 0, slope = 1))+ggtitle("abline")
g2 <- b + geom_hline(aes(yintercept = lat))+ggtitle("hline")
g3 <- b + geom_vline(aes(xintercept = long))+ggtitle("vline")
g4 <- b + geom_segment(aes(yend = lat + 1, xend = long + 1))+ggtitle("segments")
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```

## ONE VARIABLE Continuous

```{r}
str(mpg$hwy)
c <- ggplot(mpg, aes(hwy)) 
c2 <- ggplot(mpg)

g1 <- c + geom_area(stat = "bin") + ggtitle("area")
# x, y, alpha, color, fill, linetype, size
g2 <- c + geom_density(kernel = "gaussian") + ggtitle("density")
# x, y, alpha, color, fill, group, linetype, size, weight
g3 <- c + geom_dotplot() + ggtitle("dots")
# x, y, alpha, color, fill
g4 <- c + geom_freqpoly() +ggtitle("freq")
# x, y, alpha, color, group, linetype, size
g5 <- c + geom_histogram(binwidth = 5) + ggtitle("histogram")
# x, y, alpha, color, fill, linetype, size, weight
g6 <- c2 + geom_qq(aes(sample = hwy)) + ggtitle("qq")
# x, y, alpha, color, fill, linetype, size, weight
gridExtra::grid.arrange(g1,g2,g3,g4,g5,g6,nrow = 2)
```


## ONE VARIABLE discrete

```{r}
d <- ggplot(mpg, aes(fl))
d + geom_bar()
# x, alpha, color, fill, linetype, size, weight
```

## TWO VARIABLES

### both continuous

```{r}
e <- ggplot(mpg, aes(cty, hwy))
g1 <- e+ geom_label(aes(label = cty), nudge_x = 1,
nudge_y = 1)  + ggtitle("label")
#- x, y, label, alpha, angle, color,
# family, fontface, hjust, lineheight, size, vjust
g2 <- e + geom_point() + ggtitle("points")
# x, y, alpha, color, fill, shape, size, stroke
g3 <- e + geom_quantile() + ggtitle("quantile")
# x, y, alpha, color, group, linetype, size, weight
g4 <- e + geom_rug(sides = "bl")+ ggtitle("rug")
 # x, y, alpha, color, linetype, size
g5 <- e + geom_smooth(method = lm) + ggtitle("smooth")
 # x, y, alpha, color, fill, group, linetype, size, weight
g6 <- e + geom_text(aes(label = cty), nudge_x = 1,
nudge_y = 1) +ggtitle("text")
# - x, y, label, alpha, angle, color,family, fontface, hjust, lineheight, size, vjust
gridExtra::grid.arrange(g1,g2,g3,g4,g5,g6,nrow = 2)
```





### two vars :one discrete, one continuous

```{r}
f <- ggplot(mpg, aes(class, hwy))
g1 <- f + geom_col() + ggtitle("col") 
#x, y, alpha, color, fill, group, linetype, size
g2 <- f + geom_boxplot() + ggtitle("boxplot")
#x, y, lower, middle, upper, ymax, ymin, alpha,
#color, fill, group, linetype, shape, size, weight
g3 <- f + geom_dotplot(binaxis = "y", stackdir = "center")+ ggtitle("dotplot")
#x, y, alpha, color, fill, group
g4 <- f + geom_violin(scale = "area")+ ggtitle("violin")
#x, y, alpha, color, fill, group, linetype, size, weight

gridExtra::grid.arrange(g1,g2,nrow = 2)
gridExtra::grid.arrange(g3,g4,nrow = 2)
```

### both discrete

```{r}
g <- ggplot(diamonds, aes(cut, color))
g1 <- g + geom_count() +ggtitle("count")
#x, y, alpha, color, fill, shape, size, stroke
g2 <- e + geom_jitter(height = 2, width = 2) + ggtitle("jitter")
#x, y, alpha, color, fill, shape, size
gridExtra::grid.arrange(g1,g2,nrow = 1)
```

```{r}
library(hexbin)
h <- ggplot(diamonds, aes(carat, price))

g1 <- h + geom_bin2d(binwidth = c(0.25, 500)) + ggtitle("bin2d")
#x, y, alpha, color, fill, linetype, size, weight
g2 <- h + geom_density_2d() + ggtitle("density_2d")
#x, y, alpha, color, group, linetype, size
g3 <- h + geom_hex()+ ggtitle("hex")
# x, y, alpha, color, fill, size
gridExtra::grid.arrange(g1,g2,g3,nrow = 2)
```

## continuous function

```{r}
i <- ggplot(economics, aes(date, unemploy))

g1 <- i + geom_area() + ggtitle("area")
#x, y, alpha, color, fill, linetype, size
g2 <- i + geom_line() + ggtitle("line")
#x, y, alpha, color, group, linetype, size
g3 <- i + geom_step(direction = "hv") + ggtitle("step")
#x, y, alpha, color, group, linetype, size
gridExtra::grid.arrange(g1,g2,g3,nrow = 2)
```

## visualizing error

```{r}
df <- data.frame(grp = c("A", "B"), fit = 4:5, se = 1:2)
j <- ggplot(df, aes(grp, fit, ymin = fit - se, ymax = fit + se))
#head(df)
#head(j)

g1 <- j + geom_crossbar(fatten = 2) +ggtitle("crossbar")
#- x, y, ymax,ymin, alpha, color, fill, group, linetype, size
g2 <- j + geom_errorbar() +ggtitle("errorbarbar")
#- x, ymax, ymin,alpha, color, group, linetype, size, width
#Also geom_errorbarh().
g3 <- j + geom_linerange() +ggtitle("linerange")
#x, ymin, ymax, alpha, color, group, linetype, size
g4 <- j + geom_pointrange()  +ggtitle("pointrange")
#- x, y, ymin, ymax,alpha, color, fill, group, linetype, shape, size
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```

## maps

```{r}
data <- data.frame(murder = USArrests$Murder,
state = tolower(rownames(USArrests)))
map <- map_data("state")
k <- ggplot(data, aes(fill = murder))
k + geom_map(aes(map_id = state), map = map)+ 
expand_limits(x = map$long, y = map$lat)
#map_id, alpha, color, fill, linetype, size
```

## THREE VARIABLES

```{r}
seals$z <- with(seals, sqrt(delta_long^2 + delta_lat^2)); 
l <- ggplot(seals, aes(long, lat))
l + geom_contour(aes(z = z))
#x, y, z, alpha, color, group, linetype, size, weight
```

```{r}
l + geom_raster(aes(fill = z), hjust = 0.5,
vjust = 0.5, interpolate = FALSE)
#x, y, alpha, fill
l + geom_tile(aes(fill = z))
# x, y, alpha, color, fill, linetype, size, width
```

## Scales

1- GENERAL PURPOSE SCALES

-   Use with most aesthetics

-   scale\_\*\_continuous() - Map cont' values to visual ones.

-   scale\_\*\_discrete() - Map discrete values to visual ones.

-   scale\_\*\_binned() - Map continuous values to discrete bins.

-   scale\_\*\_identity() - Use data values as visual ones.

-   scale\_\*\_manual(values = c()) - Map discrete values tomanually chosen visual ones.

-   scale\_\*\_date(date_labels = "%m/%d"),

-   date_breaks = "2 weeks") - Treat data values as dates.

-   scale\_\*\_datetime() - Treat data values as date times.

-   Same as scale\_\*\_date(). See ?strptime for label formats.

**X & Y LOCATION SCALES**

Use with x or y aesthetics (x shown here)

-   scale_x\_log10() - Plot x on log10 scale.

-   scale_x\_reverse() - Reverse the direction of the x axis.

-   scale_x\_sqrt() - Plot x on square root scale.

```{r}
n <- d + geom_bar(aes(fill = fl))
n + scale_fill_manual(
values = c("skyblue", "royalblue", "blue", "navy"),
limits = c("d", "e", "p", "r"), breaks =c("d", "e", "p", "r"),
name = "fuel", labels = c("D", "E", "P", "R"))
```

```{r}
n + scale_fill_brewer(palette = "Blues")
#For palette choices:RColorBrewer::display.brewer.all()
n + scale_fill_grey(start = 0.2,
end = 0.8, na.value = "red")
```

## COLOR AND FILL SCALES (CONTINUOUS)

```{r}
o <- c + geom_dotplot(aes(fill = ..x..))
g1 <- o + scale_fill_distiller(palette = "Blues") + ggtitle("distiller")
g2 <- o + scale_fill_gradient(low="red", high="yellow") + ggtitle("gradient")
g3 <- o + scale_fill_gradient2(low = "red", high = "blue",
mid = "white", midpoint = 25)+ ggtitle("gradient2")
g4 <- o + scale_fill_gradientn(colors = topo.colors(6)) + ggtitle("gradientn")
#Also: rainbow(), heat.colors(), terrain.colors(),
#cm.colors(), RColorBrewer::brewer.pal()
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```

## SHAPE AND SIZE SCALES

```{r}
p <- e + geom_point(aes(shape = fl, size = cyl))
g1 <- p + scale_shape() + scale_size() + ggtitle("shape & size")
g2 <- p + scale_shape_manual(values = c(3:7))+ ggtitle("shape_manual")

g3 <- p + scale_radius(range = c(1,6)) + ggtitle("radius")
g4 <- p + scale_size_area(max_size = 6) + ggtitle("size_area")
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```

## Coordinate Systems

```{r}
r <- d + geom_bar()
g1 <- r + coord_cartesian(xlim = c(0, 5)) + ggtitle("cartesian")
# - xlim, ylim The default cartesian coordinate system.
g2 <- r + coord_fixed(ratio = 1/3) + ggtitle("fixed")
# ratio, xlim, ylim - Cartesian coordinates with
#fixed aspect ratio between x and y units.
g3 <- ggplot(mpg, aes(y = fl)) + geom_bar()
#Flip cartesian coordinates by switching
#x and y aesthetic mappings.

g4 <- r + coord_trans(y = "sqrt") + ggtitle("trans")
#- x, y, xlim, ylim
#Transformed cartesian coordinates. Set xtrans
#and ytrans to the name of a window function.
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```
```{r}
r + coord_polar(theta = "x", direction=1)
#theta, start, direction - Polar coordinates.
```


## Position Adjustments

```{r}
s <- ggplot(mpg, aes(fl, fill = drv))
g1 <- s + geom_bar(position = "dodge") + ggtitle("dodge")
# Arrange elements side by side.

g2 <- s + geom_bar(position = "fill") + ggtitle("fill")
g3 <- e + geom_point(position = "jitter") + ggtitle("jitter")
#e + geom_label(position = "nudge")
g4 <- s + geom_bar(position = "stack") + ggtitle("stack")
gridExtra::grid.arrange(g1,g2,g3,g4,nrow = 2)
```

## Themes

```{r}
g1 <-  r + theme_bw()
g2 <-  r + theme_gray()
g3 <-  r + theme_dark()
g4 <-  r + theme_classic()
g5 <-  r + theme_light()
g6 <-  r + theme_linedraw()
g7 <-  r + theme_minimal()

gridExtra::grid.arrange(g1,g2,g3,g4,g5,g6,g7,nrow = 2)
```

## Faceting

```{r}
t <- ggplot(mpg, aes(cty, hwy)) + geom_point()
t + facet_grid(cols = vars(fl))
# Facet into columns based on fl.
t + facet_grid(rows = vars(year))
# Facet into rows based on year.
t + facet_grid(rows = vars(year), cols = vars(fl))
# Facet into both rows and columns.
t + facet_wrap(vars(fl))
# Wrap facets into a rectangular layout
```

```{r}
t + facet_grid(rows = vars(drv), cols = vars(fl),
scales = "free") 
```

## Labels and Legends

```{r}
#Use labs() to label the elements of your plot.
t + labs(x = "New x axis label", y = "New y axis label",
title ="Add a title above the plot",
subtitle = "Add a subtitle below title",
caption = "Add a caption below plot",
alt = "Add alt text to the plot",
#<aes> = "New <aes> legend title"
        )
t + annotate(geom = "text", x = 8, y = 9, label = "A")
#Places a geom with manually selected aesthetics.
```

```{r}
p + guides(x = guide_axis(n.dodge = 2)) 
#Avoid crowdedor overlapping labels with guide_axis(n.dodge or angle).
n + guides(fill = "none") 
# Set legend type for eachaesthetic: colorbar, legend, or none (no legend).
```

```{r}
n + theme(legend.position = "bottom")
#Place legend at "bottom", "top", "left", or “right”.
n + scale_fill_discrete(name = "Title",
labels = c("A", "B", "C", "D", "E"))
#Set legend title and labels with a scale function.
```

```{r}
# Without clipping (preferred):
t + coord_cartesian(xlim = c(0, 100), ylim = c(10, 20))
# With clipping (removes unseen data points):
t + xlim(0, 100) + ylim(10, 20)
t + scale_x_continuous(limits = c(0, 100)) +
scale_y_continuous(limits = c(0, 100))
```
